---
title: "02Demographics"
format: html
editor: visual
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set( echo = FALSE, # hide code by default 
                        message = FALSE, # hide messages 
                        warning = FALSE, # hide warnings 
                        results = "hide" # hide printed output from code 
                        )

```

# Demographics

Prep work space and load packages

```{r}
source("00_setup.R")
```

Load CHILDES transcripts/ utterance dataframe in English-North America

```{r,echo=TRUE, results="show"}
d_eng_na  <- readRDS(here("01Data", "01Childes", "d_eng_na_transcripts.rds"))
ut_eng_na <- readRDS(here("01Data", "01Childes", "ut_eng_na_utterances.rds"))
head(ut_eng_na)

```

Load participant information: Number of Transcripts, transcripts, utterances by roles

```{r,echo=TRUE, results="show"}
count_transcripts <- nrow(d_eng_na)
count_transcripts

count_by_role <- ut_eng_na %>%
  group_by(speaker_role) %>%
  summarise(
  n_speakers   = n_distinct(speaker_id),
  n_transcripts = n_distinct(transcript_id),
  n_utterances = n(),
    .groups = "drop"
  ) 
#Order by number of transcripts
count_by_role <- count_by_role %>%
  arrange(desc(n_transcripts)) %>%
  mutate(
    speaker_role = factor(speaker_role,
                          levels = speaker_role)
  )

count_by_role

count_by_role_exc_child <- count_by_role %>% 
  filter(speaker_role != "Target_Child")

count_by_role_exc_child
```

Pie Chart showing number of transcripts by role excluding target child

```{r}
p_role_dist <- ggplot(count_by_role_exc_child,
       aes(x = "", y = n_transcripts, fill = speaker_role)) +
  geom_col(width = 1) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Distribution of Transcripts by Role")

p_role_dist

ggsave(
  here("03Output", "fig_role_distribution.png"),
  plot = p_role_dist,
  width = 6,
  height = 6,
  dpi = 300
)
```

### Group conversational partners

Grouping based on Bronfenbrenner’s ecological systems

Group 1: Mother (MOT)

Group 2: Father (FAT)

Group 3: Adult with existing relationship (ADT): Grandmother, Grandfather, Relative, Caretaker, Caregiver, Teacher

Group 4: Siblings (SIB): Sister, Brother, Sibling

Group 5: Other children (OTH_C): Child, Friend, Playmate, Student, Girl, Teenager

Group 6: Other adults (Exosystem) (OTH_A) : Investigator, Media, Unidentified, Adult, Visitor, Participants, Environment, Male, Uncertain

```{r}
source(here("02Script", "03RoleMap.R"))

count_by_role_exc_child <- count_by_role_exc_child %>%
  left_join(role_category_map, by = "speaker_role")
count_by_role <- count_by_role %>% 
  left_join(role_category_map, by = "speaker_role")

count_by_role_exc_child

#Sanity check 
count_by_role_exc_child %>%
  filter(is.na(role_category)) %>%
  distinct(speaker_role)
```

Number of participant, transcripts, utterance by each category

```{r}
count_by_category <- count_by_role_exc_child %>%
  group_by(role_category) %>%
  summarise(
    n_speakers    = sum(n_speakers),
    n_transcripts = sum(n_transcripts),
    n_utterances  = sum(n_utterances),
    .groups = "drop"
  ) %>%
  arrange(desc(n_transcripts))

count_by_category
```

Load information on the target child: number of target child, sex, age

```{r,echo=TRUE, results="show"}
d_target_child <- readRDS(here("01Data", "01Childes", "d_target_child.rds"))

sum_sex <- d_target_child %>%
  count(sex)
sum_sex

sum_age <-d_target_child %>%
  summarise(
    mean_age = mean(min_age, na.rm = TRUE),
    sd_age = sd(min_age, na.rm = TRUE),
    min_age = min(min_age, na.rm = TRUE),
    max_age = max(max_age, na.rm = TRUE)
  )
sum_age
```

Load information about the SES and education

```{r,echo=TRUE, results="show"}
sum_ses <- d_target_child %>%
  count(ses)
sum_ses

sum_education <-d_target_child %>%
  count(education)
sum_education
```
