---
title: "06RQ1Pipeline"
output: html_document
execute:
  dir: project

---

```{r setup, include=FALSE}
 knitr::opts_chunk$set( echo = FALSE, # hide code by default 
                        message = FALSE, # hide messages 
                        warning = FALSE, # hide warnings 
                        results = "hide" # hide printed output from code 
                        )

```

Prep work space and load packages

```{r}
source(here::here("00_setup.R"))
```

Load CHILDES transcripts/ utterance dataframe in English-North America

```{r,echo=TRUE, results="show"}
d_eng_na  <- readRDS(here("01Data", "01Childes", "d_eng_na_transcripts.rds"))
ut_eng_na <- readRDS(here("01Data", "01Childes", "ut_eng_na_utterances.rds"))
utt_partners_emotions <- readRDS(here("01Data", "02Derived", "utt_partners_emotions.rds"))

file_lex <- read.csv("~/Desktop/R_WD/ThesisDataAnalysis/1987-Affectivelexicon-foundations_words.csv",
                     stringsAsFactors = FALSE)
emotion_words <-file_lex$Word
```

# RQ1 Dataframe Prep

Identify if any transcripts, only the target child is speaking (no conversational partner speech)

```{r}
# Identify only-child transcripts
onlychild_ids <- ut_eng_na %>%
  group_by(transcript_id) %>%
  summarise(
    only_child = all(na.omit(speaker_code) == "CHI"),
    .groups = "drop"
  ) %>%
  filter(only_child) %>%
  pull(transcript_id)

#sanity check
length(onlychild_ids)               # how many such transcripts


#transcript:
eng_na_rq1 <- d_eng_na %>%
  filter(!transcript_id %in% onlychild_ids)
count_transcripts_rq1 <- nrow(eng_na_rq1)

count_transcripts_rq1
#number of transcripts where there is a conversational partner
```

Partner count

```{r}

partner_summary_table <- utt_partners_emotions %>%
  # Step 1: count unique partner categories per transcript
  group_by(transcript_id) %>%
  summarise(
    n_partner_categories = n_distinct(role_category, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  # Step 2: count how many transcripts fall into each category count
  count(n_partner_categories, name = "n_transcripts") %>%
  arrange(n_partner_categories)

partner_summary_table

```

Across all transcripts, how many times each words are used + frequency

```{r}

total_word_count <- utt_partners_emotions %>%
  summarise(
    total_words = sum(str_count(gloss_lower, "\\b\\w+\\b"))
  ) %>%
  pull(total_words)

total_emotion_words <- utt_partners_emotions %>%
  summarise(
    across(all_of(emotion_words), ~ sum(.x, na.rm = TRUE)),
    emotion_count = sum(emotion_count, na.rm = TRUE),
    .groups = "drop"
  )

emotion_counts_long <- total_emotion_words %>%
  pivot_longer(
    cols = everything(),
    names_to = "word",
    values_to = "count"
  ) %>%
  mutate(
    freq_per_1000_words = (count / total_word_count) * 1000
  ) %>%
  arrange(desc(freq_per_1000_words))

write.csv(
  emotion_counts_long,
  file = here("03Output", "emotion word count and frequency.csv"),
  row.names = FALSE
)


emotion_counts_long
```

By Role

```{r}
total_emotion_words_by_role <- utt_partners_emotions %>%
  group_by(role_category) %>%
  summarise(
    across(all_of(emotion_words), ~ sum(.x, na.rm = TRUE)),
    emotion_count = sum(emotion_count, na.rm = TRUE),
    .groups = "drop"
  )

total_word_count_by_role <- utt_partners_emotions %>%
  mutate(word_count = str_count(gloss_lower, "\\b\\w+\\b")) %>%
  group_by(role_category) %>%
  summarise(
    total_words = sum(word_count, na.rm = TRUE),
    .groups = "drop"
  )

emotion_counts_long_by_role <- total_emotion_words_by_role %>%
  left_join(total_word_count_by_role, by = "role_category") %>%
  pivot_longer(
    cols = -c(role_category, total_words),
    names_to = "word",
    values_to = "count"
  ) %>%
  mutate(
    freq_per_1000_words = (count / total_words) * 1000
  ) %>%
  arrange(role_category, desc(freq_per_1000_words))

total_emotion_counts_long_by_role <- emotion_counts_long_by_role %>%
  filter(
    word == "emotion_count"
  )
total_emotion_counts_long_by_role

write.csv(
  emotion_counts_long_by_role,
  file = here("03Output", "emotion_word_count_and_frequency_by_role.csv"),
  row.names = FALSE
)
```

## Problem 1. Disambiguation. e.g. like, well, kind.

### Step 1: Flag ambiguous words from main emotion word list


get a list of emotion word count \>10

```{r}
emotion_words_gt10 <- emotion_counts_long %>%
  filter(count > 10) %>%
  pull(word)
emotion_words_gt10
```

Sample randomly up to 30 example utterance each word, if \<30, print all utterances

```{r}

examples_word <- map_dfr(emotion_words_gt10, function(w) {
  
  # all utterances where this word appears at least once
  df_w <- utt_partners_emotions %>%
    filter(.data[[w]] > 0)   # use the column named w
  
  # if no utterances contain this word, skip it
  if (nrow(df_w) == 0) return(NULL)
  
  # how many to sample (max 20, but not more than we have)
  n_samp <- min(20L, nrow(df_w))
  
  df_w %>%
    slice_sample(n = n_samp) %>%
    mutate(target_word = w)
}) 

write.csv(
  examples_word,
  here("01Data","02Derived","gt10words_20examples.csv")
  ,
  row.names = FALSE
)
```

### Step 2 :Subset utterances containing those words

```{r}
ambiguous_cols <- c(
  #POS tagging
  "like","well","kind","fine","blue","merry","moved","certain","high","touched","patient",
  
  #S-BERT
  "lost","quiet",
  "low",
  
  #Manual
  "rotten","pride","ill","tender","crushed","gloomy","faint","odd")
ambiguous_cols


ambiguous_word_counts <- emotion_counts_long %>%
  filter(word %in% ambiguous_cols) %>%
  select(word, count) %>%
  arrange(desc(count))

ambiguous_word_counts
amb_utt <- utt_partners_emotions %>%
  filter(if_any(all_of(ambiguous_cols), ~ .x > 0)) %>% mutate(
    ambiguous_word = apply(
      select(., all_of(ambiguous_cols)),
      1,
      function(x) {
        paste(ambiguous_cols[x > 0], collapse = "|")
      }
    )
  ) %>% 
  select(ambiguous_word, id, gloss,role_category,gloss_lower)
head(amb_utt)

#sanity check
amb_utt %>%
  count(ambiguous_word, sort = TRUE)

saveRDS(amb_utt, here("01Data", "02Derived", "amb_utterances.rds"))

```

### Step 3: POS and dependency-based disambiguation (spaCy)

Data Prep go to 05SpacyRTagging_RQ1.R

```{r}
amb_utt <- readRDS(here("01Data", "02Derived", "amb_utterances.rds"))
```

#### 3.1 Like

new dataframe containing only like-included utterances + original utterance id.

```{r}
df_like <- readRDS(here("01Data", "02Derived","01Dataframe", "df_like_utterances.rds"))
tok_like <- readRDS(here("01Data", "02Derived", "02Spacy Parsed","spacy_like_parsed.rds"))
```

define the rules for like: POS = verb + exclude "look(s) like“

```{r}
# 1) Like tokens (token-level)
like_tok <- tok_like %>%
  filter(lemma == "like") %>%
  mutate(doc_id = as.character(doc_id)) %>%
  select(doc_id, token_id, lemma, pos, dep_rel, head_token_id)

# 2) Head token lookup (for excluding "look like")
heads <- tok_like %>%
  mutate(doc_id = as.character(doc_id)) %>%
  select(doc_id, token_id, head_token = token)

# 3) Preference pattern at doc level (counts ALL like tokens in that utterance as emotion)
pref_docs <- amb_utt %>%
  transmute(doc_id = as.character(id), text = gloss_lower) %>%
  filter(
    str_detect(text, "\\b(would|could|can|do|did)\\s+you\\s+like\\b") |
      str_detect(text, "\\b(i\\s*'d|i\\s+would|i\\s+could)\\s+like\\b")
  ) %>%
  distinct(doc_id) %>%
  mutate(pref_rule = TRUE)

# 4) Token-level emotion rule + aggregate counts per doc
like_counts_by_doc <- like_tok %>%
  left_join(heads, by = c("doc_id", "head_token_id" = "token_id")) %>%
  mutate(
    is_look_like = head_token %in% c("look", "looks", "looked", "looking"),
    like_token_is_emotion = (pos == "VERB") & !is_look_like
  ) %>%
  left_join(pref_docs, by = "doc_id") %>%
  mutate(
    pref_rule = if_else(is.na(pref_rule), FALSE, pref_rule),
    like_token_is_emotion = like_token_is_emotion | pref_rule
  ) %>%
  group_by(doc_id) %>%
  summarise(
    n_like_tokens = n(),
    n_like_emotion = sum(like_token_is_emotion),
    n_like_nonemotion = n_like_tokens - n_like_emotion,
    like_pos = paste(sort(unique(pos)), collapse = "|"),
    .groups = "drop"
  )

saveRDS(
  like_counts_by_doc,
  here("01Data", "02Derived","03Annotation", "like_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

like_counts_by_doc <- readRDS(
  here("01Data", "02Derived","03Annotation","like_counts_by_doc.rds")
)


amb_utt_like_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(like_counts_by_doc, by = "doc_id") %>%
  mutate(
    n_like_tokens = coalesce(n_like_tokens, 0L),
    n_like_emotion = coalesce(n_like_emotion, 0L),
    n_like_nonemotion = coalesce(n_like_nonemotion, 0L)
  )

```

Count Check

```{r}
#token level inclusion and exclusion
like_totals<- amb_utt_like_flagged %>%
  summarise(
    # token level
    total_like_tokens = sum(n_like_tokens, na.rm = TRUE),
    total_like_emotion_tokens = sum(n_like_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_like = sum(n_like_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_like = sum(n_like_emotion > 0, na.rm = TRUE)
  )

like_totals

#how many utterance with more than two likes 
amb_utt_like_flagged %>%
  summarise(
    utt_with_2plus_likes = sum(n_like_emotion >= 2, na.rm = TRUE),
    utt_with_3plus_likes = sum(n_like_emotion >= 3, na.rm = TRUE),
    max_likes_in_utt     = max(n_like_emotion, na.rm = TRUE)
  )

#check emotion inclusion examples

set.seed(1)

amb_utt_like_flagged %>%
  filter(n_like_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_like_flagged %>%
  filter(n_like_tokens > 0, n_like_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_like_flagged %>%
  filter(id == 2613441)
```

how words are used in speech common error: emotion word used as ADJ but not expressing emotion high blue + noun

#### 3.2 Well

rules: well as adj

```{r}
#prepare dataset
df_well <- readRDS(here("01Data", "02Derived",  "01Dataframe","df_well_utterances.rds"))
tok_well <- readRDS(here("01Data", "02Derived", "02Spacy Parsed","spacy_well_parsed.rds"))
```

POS tagging and rule applying

```{r}
# 0) Build a token count per utterance (doc)
doc_n_tokens <- tok_well %>%
  mutate(doc_id = as.character(doc_id)) %>%
  count(doc_id, name = "n_tokens")

# 1) Pull WELL tokens from spacy output
well_tok <- tok_well %>%
  mutate(doc_id = as.character(doc_id)) %>%
  filter(lemma == "well") %>%
  select(doc_id, token_id, lemma, pos, dep_rel, head_token_id, token)

# 2) Add utterance text (for regex exclusions)
utt_text <- amb_utt %>%
  transmute(doc_id = as.character(id), text = gloss_lower)

# 3) Identify which *tokens* of "well" are emotion-uses
#    (token-level, so we can count >1 per utterance)
well_emotion_tokens <- well_tok %>%
  left_join(doc_n_tokens, by = "doc_id") %>%
  left_join(utt_text, by = "doc_id") %>%
  filter(
    pos %in% c("ADJ", "ADV"),            # keep ADJ/ADV
    n_tokens > 1,                        # exclude single-word utterances
    !str_detect(text, "^\\s*well\\b"),   # exclude sentence-initial discourse "well ..."
    !str_detect(text, "\\bas\\s+well\\b")# exclude "as well"
  ) %>%
  distinct(doc_id, token_id)

# 4) Count WELL tokens per utterance (all, emotion, non-emotion)
well_counts_by_doc <- well_tok %>%
  distinct(doc_id, token_id) %>%  # avoid accidental duplicates
  left_join(
    well_emotion_tokens %>% mutate(is_well_emotion = TRUE),
    by = c("doc_id", "token_id")
  ) %>%
  mutate(is_well_emotion = if_else(is.na(is_well_emotion), FALSE, is_well_emotion)) %>%
  group_by(doc_id) %>%
  summarise(
    n_well_tokens     = n(),
    n_well_emotion    = sum(is_well_emotion),
    n_well_nonemotion = n_well_tokens - n_well_emotion,
    .groups = "drop"
  )

saveRDS(
  well_counts_by_doc,
  here("01Data", "02Derived","03Annotation", "well_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

well_counts_by_doc <- readRDS(
  here("01Data", "02Derived","03Annotation","well_counts_by_doc.rds")
)

amb_utt_well_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(well_counts_by_doc, by = "doc_id") %>%
  mutate(
    n_well_tokens     = coalesce(n_well_tokens, 0L),
    n_well_emotion    = coalesce(n_well_emotion, 0L),
    n_well_nonemotion = coalesce(n_well_nonemotion, 0L)
  )

```

Count Check

```{r}
#token level inclusion and exclusion
well_totals<- amb_utt_well_flagged %>%
  summarise(
    # token level
    total_well_tokens = sum(n_well_tokens, na.rm = TRUE),
    total_well_emotion_tokens = sum(n_well_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_well = sum(n_well_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_well = sum(n_well_emotion > 0, na.rm = TRUE)
  )

well_totals

#how many utterance with more than two likes 
amb_utt_well_flagged %>%
  summarise(
    utt_with_2plus_wells = sum(n_well_emotion >= 2, na.rm = TRUE),
    utt_with_3plus_wells= sum(n_well_emotion >= 3, na.rm = TRUE),
    max_well_in_utt     = max(n_well_emotion, na.rm = TRUE)
  )

#check emotion inclusion examples

set.seed(1)

amb_utt_well_flagged %>%
  filter(n_well_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_well_flagged %>%
  filter(n_well_tokens > 0, n_well_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_well_flagged %>%
  filter(id == 17692225)
```

#### 3.3. Kind

rule: keep ADJ + exclude "that kind (of thing)"

```{r}
#prepare dataset
df_kind <- readRDS(here("01Data", "02Derived","01Dataframe", "df_kind_utterances.rds"))
tok_kind <- readRDS(here("01Data", "02Derived", "02Spacy Parsed","spacy_kind_parsed.rds"))

```

```{r}
#1) Pull KIND tokens (token-level) ----
kind_tok <- tok_kind %>%
  mutate(doc_id = as.character(doc_id)) %>%
  filter(lemma == "kind") %>%
  select(doc_id, token_id, lemma, pos, dep_rel, head_token_id, token)

kind_tok_w_text <- kind_tok %>%
  left_join(df_kind, by = "doc_id")   # adds text column for exclusions

# ---- 2) Token-level emotion KIND (can count >1 per utterance) ----
kind_emotion_tokens <- kind_tok_w_text %>%
  filter(pos == "ADJ") %>%
  # exclude demonstrative/question "kind" noun uses: "this kind", "what kind", etc.
  filter(!str_detect(text, "\\b(this|that|these|those|what|which)\\s+kind\\b")) %>%
  distinct(doc_id, token_id)

# ---- 3) Aggregate counts per utterance ----
kind_counts_by_doc <- kind_tok_w_text %>%
  distinct(doc_id, token_id) %>%  # guard against accidental duplicates
  left_join(
    kind_emotion_tokens %>% mutate(is_kind_emotion = TRUE),
    by = c("doc_id", "token_id")
  ) %>%
  mutate(is_kind_emotion = if_else(is.na(is_kind_emotion), FALSE, is_kind_emotion)) %>%
  group_by(doc_id) %>%
  summarise(
    n_kind_tokens     = n(),
    n_kind_emotion    = sum(is_kind_emotion),
    n_kind_nonemotion = n_kind_tokens - n_kind_emotion,
    .groups = "drop"
  )

saveRDS(
  kind_counts_by_doc,
  here("01Data", "02Derived","03Annotation", "kind_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

kind_counts_by_doc <- readRDS(
  here("01Data", "02Derived","03Annotation","kind_counts_by_doc.rds")
)

amb_utt_kind_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(kind_counts_by_doc, by = "doc_id") %>%
  mutate(
    n_kind_tokens     = coalesce(n_kind_tokens, 0L),
    n_kind_emotion    = coalesce(n_kind_emotion, 0L),
    n_kind_nonemotion = coalesce(n_kind_nonemotion, 0L)
  )

```

```{r}

#token level inclusion and exclusion
kind_totals <- amb_utt_kind_flagged %>%
  summarise(
    # token level
    total_kind_tokens = sum(n_kind_tokens, na.rm = TRUE),
    total_kind_emotion_tokens = sum(n_kind_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_kind = sum(n_kind_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_kind = sum(n_kind_emotion > 0, na.rm = TRUE)
  )

kind_totals

#how many utterance with more than two likes 
amb_utt_kind_flagged %>%
  summarise(
    utt_with_2plus_kinds = sum(n_kind_emotion >= 2, na.rm = TRUE),
    utt_with_3plus_kinds= sum(n_kind_emotion >= 3, na.rm = TRUE),
    max_kind_in_utt     = max(n_kind_emotion, na.rm = TRUE)
  )

#check emotion inclusion examples

set.seed(1)

amb_utt_kind_flagged %>%
  filter(n_kind_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_kind_flagged %>%
  filter(n_kind_tokens > 0, n_kind_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_kind_flagged %>%
  filter(id == 17692225)
```

####3.4 Fine

```{r}
df_fine <- readRDS(here("01Data", "02Derived", "01Dataframe","df_fine_utterances.rds"))
tok_fine <- readRDS(here("01Data", "02Derived","02Spacy Parsed", "spacy_fine_parsed.rds"))

```

```{r}
tok_fine2 <- tok_fine %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id)
  )

fine_tok <- tok_fine2 %>%
  filter(lemma == "fine") %>%
  select(doc_id, token_id, lemma, pos)

next_tok <- tok_fine2 %>%
  select(doc_id, token_id, next_pos = pos, next_token = token)

# join each "kind" token to its NEXT token in the same doc
fine_tok_with_next <- fine_tok %>%
  left_join(
    next_tok %>% rename(next_token_id = token_id),
    by = "doc_id"
  ) %>%
  filter(next_token_id == token_id + 1) %>%
  select(-next_token_id)

# 1) Token-level exclusion: "fine" as ADJ followed by NOUN/PROPN (a fine lady)
fine_excluded_tokens_adj_noun <- fine_tok_with_next %>%
  filter(pos == "ADJ", next_pos %in% c("NOUN", "PROPN")) %>%
  distinct(doc_id, token_id)

# 2) Doc-level exclusion: discourse-initial / standalone "fine"
fine_exclude_start_standalone_docs <- df_fine %>%
  filter(
    # starts with "fine"
    str_detect(text, "^\\s*fine\\b") &
    
    # standalone / discourse use: punctuation or end
    str_detect(text, "^\\s*fine(\\s*[\\.,—-]|\\s*$)") &
    
    # NOT emotional constructions
    !str_detect(
      text,
      "^\\s*(it|that|you)\\s*(is|'s|are|'re)?\\s+fine\\b"
    )
  ) %>%
  distinct(doc_id)


# 3) Combine BOTH token-level exclusion sets
fine_excluded_tokens_all <- bind_rows(
  fine_excluded_tokens_adj_noun,
  fine_exclude_start_standalone_docs
) %>%
  distinct(doc_id, token_id)

fine_all_counts <- fine_tok %>%
  count(doc_id, name = "n_fine_tokens")

fine_excluded_counts <- fine_excluded_tokens_all %>%
  count(doc_id, name = "n_fine_nonemotion")


fine_emotion_counts <- fine_all_counts %>%
  left_join(fine_excluded_counts, by = "doc_id") %>%
  mutate(
    n_fine_nonemotion = coalesce(n_fine_nonemotion, 0L),
    n_fine_emotion = n_fine_tokens - n_fine_nonemotion
  ) %>%
  select(doc_id, n_fine_tokens, n_fine_emotion, n_fine_nonemotion)

saveRDS(
  fine_emotion_counts,
  here("01Data", "02Derived","03Annotation", "fine_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

fine_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","fine_counts_by_doc.rds")
)

amb_utt_fine_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(fine_emotion_counts, by = "doc_id") %>%
  mutate(
    n_fine_tokens = coalesce(n_fine_tokens, 0L),
    n_fine_emotion = coalesce(n_fine_emotion, 0L),
    n_fine_nonemotion = coalesce(n_fine_nonemotion, 0L)
  )

```

```{r}
fine_totals <- amb_utt_fine_flagged %>%
  summarise(
    # token level
    total_fine_tokens = sum(n_fine_tokens, na.rm = TRUE),
    total_fine_emotion_tokens = sum(n_fine_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_fine = sum(n_fine_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_fine = sum(n_fine_emotion > 0, na.rm = TRUE)
  )

fine_totals

#how many utterance with more than two likes 
amb_utt_fine_flagged %>%
  summarise(
    utt_with_2plus_fines = sum(n_fine_emotion >= 2, na.rm = TRUE),
    utt_with_3plus_fines= sum(n_fine_emotion >= 3, na.rm = TRUE),
    max_fine_in_utt     = max(n_fine_emotion, na.rm = TRUE)
  )

#check emotion inclusion examples

set.seed(1)

amb_utt_fine_flagged %>%
  filter(n_fine_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_fine_flagged %>%
  filter(n_fine_tokens > 0, n_fine_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_fine_flagged %>%
  filter(id == 318002)

```

####3.5 Blue

```{r}
df_blue <- readRDS(here("01Data", "02Derived", "01Dataframe","df_blue_utterances.rds"))
tok_blue <- readRDS(here("01Data", "02Derived","02Spacy Parsed", "spacy_blue_parsed.rds"))
```

```{r}

# Make sure token_id is numeric
tok_blue2 <- tok_blue %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id)
  )

blue_tok <- tok_blue2 %>%
  filter(lemma == "blue") %>%
  select(doc_id, token_id, token, lemma, pos, dep_rel)

tok_next <- tok_blue2 %>%
  select(doc_id, token_id, next_token = token, next_lemma = lemma, next_pos = pos)

blue_tok_with_next <- blue_tok %>%
  mutate(next_token_id = token_id + 1L) %>%     # <-- compute next id
  left_join(
    tok_next,
    by = c("doc_id" = "doc_id", "next_token_id" = "token_id")
  ) %>%
  select(-next_token_id)

utt_lengths <- blue_tok_with_next %>%
  distinct(doc_id, token_id) %>%
  group_by(doc_id) %>%
  summarise(n_tokens = n(), .groups = "drop")

blue_tok_with_next <- blue_tok_with_next %>%
  left_join(utt_lengths, by = "doc_id")
blue_excluded_tokens <- blue_tok_with_next %>%
  filter(
    # 1) blue as adjective modifying a noun/proper noun
    (pos == "ADJ" & next_pos %in% c("NOUN", "PROPN")) |
      
    # 2) blue used as a noun
    (pos == "NOUN") 
  ) %>%
  distinct(doc_id, token_id)


blue_all_counts <- blue_tok %>%
  count(doc_id, name = "n_blue_tokens")

# excluded (non-emotion) "blue" tokens per utterance
blue_excluded_counts <- blue_excluded_tokens %>%
  count(doc_id, name = "n_blue_nonemotion")

# emotion counts = total - excluded
blue_emotion_counts <- blue_all_counts %>%
  left_join(blue_excluded_counts, by = "doc_id") %>%
  mutate(
    n_blue_nonemotion = coalesce(n_blue_nonemotion, 0L),
    n_blue_emotion = n_blue_tokens - n_blue_nonemotion
  ) %>%
  select(doc_id, n_blue_tokens, n_blue_emotion, n_blue_nonemotion)

saveRDS(
  blue_emotion_counts,
  here("01Data", "02Derived","03Annotation", "blue_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

blue_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","blue_counts_by_doc.rds")
)


amb_utt_blue_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(blue_emotion_counts, by = "doc_id") %>%
  mutate(
    n_blue_tokens = coalesce(n_blue_tokens, 0L),
    n_blue_emotion = coalesce(n_blue_emotion, 0L),
    n_blue_nonemotion = coalesce(n_blue_nonemotion, 0L)
  )

```

Count Check

```{r}
blue_totals <- amb_utt_blue_flagged %>%
  summarise(
    # token level
    total_blue_tokens = sum(n_blue_tokens, na.rm = TRUE),
    total_blue_emotion_tokens = sum(n_blue_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_blue = sum(n_blue_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_blue = sum(n_blue_emotion > 0, na.rm = TRUE)
  )

blue_totals


#check emotion inclusion examples

set.seed(1)

amb_utt_blue_flagged %>%
  filter(n_blue_emotion > 0) %>%
  slice_sample(n = 169) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_blue_flagged %>%
  filter(n_blue_tokens > 0, n_blue_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_blue_flagged %>%
  filter(id == 1874322)
```
#### 3.12 Merry

```{r} 
df_merry <- readRDS(here("01Data", "02Derived", "01Dataframe","df_merry_utterances.rds")) 
tok_merry <- readRDS(here("01Data", "02Derived","02Spacy Parsed", "spacy_merry_parsed.rds"))
```

Rule: filter Merry-go-around

```{r}
df_merry2 <- df_merry %>%
  mutate(
    doc_id = as.character(doc_id),
    text_lower = str_to_lower(text)
  )

tok_merry2 <- tok_merry %>%
  mutate(
    doc_id = as.character(doc_id),
    lemma_lower = str_to_lower(lemma),
    token_lower = str_to_lower(token)
  )

# --- Rule 1: Merry-go(-round/ride) family (doc-level) ---
merry_exclude_go_family_docs <- df_merry2 %>%
  filter(
    str_detect(text_lower, "\\bmerry\\s*[-–—]?\\s*go\\b") |
      str_detect(text_lower, "\\bmerry\\s*[-–—]?\\s*go\\s*[-–—]?\\s*(round|ride)\\b")
  ) %>%
  distinct(doc_id)

# tokens that are "merry"
merry_tok <- tok_merry2 %>%
  filter(lemma_lower == "merry") %>%
  select(doc_id, token_id, token, lemma, pos, lemma_lower)

merry_excluded_go_family_tokens <- merry_tok %>%
  semi_join(merry_exclude_go_family_docs, by = "doc_id") %>%
  distinct(doc_id, token_id)

# --- Rule 2: PROPN exclusion but keep Merry Christmas (doc-level guard) ---
merry_exclude_propn_not_christmas_docs <- df_merry2 %>%
  filter(
    str_detect(text_lower, "\\bmerry\\b"),
    !str_detect(text_lower, "\\bmerry\\s*[-–—]?\\s*christmas\\b")
  ) %>%
  distinct(doc_id)

merry_excluded_propn_tokens <- merry_tok %>%
  filter(pos == "PROPN") %>%
  semi_join(merry_exclude_propn_not_christmas_docs, by = "doc_id") %>%
  distinct(doc_id, token_id)

# --- Combine exclusions ---
merry_excluded_tokens_all <- bind_rows(
  merry_excluded_propn_tokens,
  merry_excluded_go_family_tokens
) %>%
  distinct(doc_id, token_id)

# --- Counts per doc ---
merry_all_counts <- merry_tok %>%
  count(doc_id, name = "n_merry_tokens")

merry_excluded_counts <- merry_excluded_tokens_all %>%
  count(doc_id, name = "n_merry_nonemotion")

merry_emotion_counts <- merry_all_counts %>%
  left_join(merry_excluded_counts, by = "doc_id") %>%
  mutate(
    n_merry_nonemotion = coalesce(n_merry_nonemotion, 0L),
    n_merry_emotion = pmax(n_merry_tokens - n_merry_nonemotion, 0L)  # safety
  ) %>%
  select(doc_id, n_merry_tokens, n_merry_emotion, n_merry_nonemotion)

saveRDS(
  merry_emotion_counts,
  here("01Data", "02Derived", "03Annotation", "merry_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

merry_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","merry_counts_by_doc.rds")
)

amb_utt_merry_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(merry_emotion_counts, by = "doc_id") %>%
  mutate(
    n_merry_tokens = coalesce(n_merry_tokens, 0L),
    n_merry_emotion = coalesce(n_merry_emotion, 0L),
    n_merry_nonemotion = coalesce(n_merry_nonemotion, 0L)
  )
merry_go_round_n_by_doc <- tok_merry2 %>%
  mutate(
    doc_id = as.character(doc_id),
    token_lower = str_to_lower(token)
  ) %>%
  filter(token_lower %in% c("merry+go+round", "merry+go+rounds")) %>%
  count(doc_id, name = "n_mgr")   # helper: how many "merry-go-round" tokens in that utterance
merry_go_round_n_by_doc

# 2) Join + shift counts: emotion -> nonemotion for those tokens
amb_utt_merry_flagged <- amb_utt_merry_flagged %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(merry_go_round_n_by_doc, by = "doc_id") %>%
  mutate(
    n_mgr = coalesce(n_mgr, 0L),

    # shift: these should NOT be emotion
    n_merry_nonemotion = n_merry_nonemotion + n_mgr,
    n_merry_emotion    = pmax(n_merry_emotion - n_mgr, 0L),
     n_merry_tokens = n_merry_emotion + n_merry_nonemotion


    # n_merry_tokens stays the same (we are reclassifying, not removing)
  ) %>%
  select(-n_mgr)

head(amb_utt_merry_flagged)
```

```{r}

#token level inclusion and exclusion
merry_totals <- amb_utt_merry_flagged %>%
  summarise(
    # token level
    total_merry_tokens = sum(n_merry_tokens, na.rm = TRUE),
    total_merry_emotion_tokens = sum(n_merry_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_merry = sum(n_merry_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_merry = sum(n_merry_emotion > 0, na.rm = TRUE)
  )

merry_totals

#how many utterance with more than two likes 
amb_utt_merry_flagged %>%
  summarise(
    utt_with_2plus_merrys = sum(n_merry_tokens >= 2, na.rm = TRUE),
    utt_with_3plus_merrys= sum(n_merry_tokens >= 3, na.rm = TRUE),
    max_merry_in_utt     = max(n_merry_tokens, na.rm = TRUE)
  )

#check emotion inclusion examples

set.seed(1)

amb_utt_merry_flagged %>%
  filter(n_merry_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_merry_flagged %>%
  filter(n_merry_tokens > 0, n_merry_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_merry_flagged %>%
  filter(id == 17486037)

```
here
#### 3.13 Moved

```{r}
df_moved <- readRDS(here("01Data", "02Derived", "01Dataframe","df_moved_utterances.rds"))
tok_moved <- readRDS(here("01Data", "02Derived","02Spacy Parsed", "spacy_moved_parsed.rds"))
```


```{r}
tok_moved2 <- tok_moved %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id),
    token_lower = str_to_lower(token),
    lemma_lower = str_to_lower(lemma)
  )
moved_tok <- tok_moved2 %>%
  filter(token_lower == "moved") %>%
  select(doc_id, token_id, token, lemma, pos)

# total moved tokens per doc
moved_all_counts <- moved_tok %>%
  count(doc_id, name = "n_moved_tokens")

# excluded = moved used as a VERB (your rule)
moved_excluded_tokens <- moved_tok %>%
  filter(pos == "VERB") %>%
  distinct(doc_id, token_id)

moved_excluded_counts <- moved_excluded_tokens %>%
  count(doc_id, name = "n_moved_nonemotion")

# derive emotion counts
moved_emotion_counts <- moved_all_counts %>%
  left_join(moved_excluded_counts, by = "doc_id") %>%
  mutate(
    n_moved_nonemotion = coalesce(n_moved_nonemotion, 0L),
    n_moved_emotion = n_moved_tokens - n_moved_nonemotion
  ) %>%
  select(doc_id, n_moved_tokens, n_moved_emotion, n_moved_nonemotion)

# save
saveRDS(
  moved_emotion_counts,
  here("01Data", "02Derived","03Annotation", "moved_counts_by_doc.rds")
)

```

```{r}

moved_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","moved_counts_by_doc.rds")
)

amb_utt_moved_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(moved_emotion_counts, by = "doc_id") %>%
  mutate(
    n_moved_tokens = coalesce(n_moved_tokens, 0L),
    n_moved_emotion = coalesce(n_moved_emotion, 0L),
    n_moved_nonemotion = coalesce(n_moved_nonemotion, 0L)
  ) %>% 
  mutate(
    n_moved_emotion = if_else(id == "600098", 0L, n_moved_emotion)
  )

head(amb_utt_moved_flagged)
```

```{r}

#token level inclusion and exclusion
moved_totals <- amb_utt_moved_flagged %>%
  summarise(
    # token level
    total_moved_tokens = sum(n_moved_tokens, na.rm = TRUE),
    total_moved_emotion_tokens = sum(n_moved_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_moved = sum(n_moved_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_moved = sum(n_moved_emotion > 0, na.rm = TRUE)
  )

moved_totals

#check emotion inclusion examples

set.seed(1)

amb_utt_moved_flagged %>%
  filter(n_moved_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_moved_flagged %>%
  filter(n_moved_tokens > 0, n_moved_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_moved_flagged %>%
  filter(id == 17486037)

```

#### 3.7 Certain
```{r}
df_certain <- readRDS(here("01Data", "02Derived", "01Dataframe","df_certain_utterances.rds"))
tok_certain <- readRDS(here("01Data", "02Derived","02Spacy Parsed", "spacy_certain_parsed.rds"))
```

```{r}
tok_certain2 <- tok_certain %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id),
    token_lower = str_to_lower(token),
    lemma_lower = str_to_lower(lemma)
  )

certain_tok <- tok_certain2 %>%
  filter(lemma_lower == "certain") %>%
  select(doc_id, token_id, token, lemma, pos)
# total moved tokens per doc
certain_all_counts <- certain_tok %>%
  count(doc_id, name = "n_certain_tokens")


tok_next <- tok_certain2 %>%
  select(
    doc_id,
    token_id,
    next_token = token,
    next_lemma = lemma,
    next_pos = pos
  )

certain_tok_with_next <- certain_tok %>%
  mutate(next_token_id = token_id + 1L) %>%
  left_join(
    tok_next,
    by = c("doc_id", "next_token_id" = "token_id")
  ) %>%
  select(-next_token_id)

certain_excluded_det_tokens <- certain_tok_with_next %>%
  filter(
    pos == "ADJ",
    next_pos %in% c("NOUN", "PROPN")
  ) %>%
  distinct(doc_id, token_id)

certain_excluded_counts <- certain_excluded_det_tokens %>%
  count(doc_id, name = "n_certain_nonemotion")

# derive emotion counts
certain_emotion_counts <- certain_all_counts %>%
  left_join(certain_excluded_counts, by = "doc_id") %>%
  mutate(
    n_certain_nonemotion = coalesce(n_certain_nonemotion, 0L),
    n_certain_emotion = n_certain_tokens - n_certain_nonemotion
  ) %>%
  select(doc_id, n_certain_tokens, n_certain_emotion, n_certain_nonemotion)

# save
saveRDS(
  certain_emotion_counts,
  here("01Data", "02Derived","03Annotation", "certain_counts_by_doc.rds")
)
```

```{r}
certain_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","certain_counts_by_doc.rds")
)

amb_utt_certain_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(certain_emotion_counts, by = "doc_id") %>%
  mutate(
    n_certain_tokens = coalesce(n_certain_tokens, 0L),
    n_certain_emotion = coalesce(n_certain_emotion, 0L),
    n_certain_nonemotion = coalesce(n_certain_nonemotion, 0L)
  )

head(amb_utt_certain_flagged)

df_certain_emotion_candidates <- amb_utt_certain_flagged %>%
  filter(n_certain_emotion > 0) %>% 
  select(
    doc_id,
    gloss = gloss,          # or your original text column
    n_certain_emotion
  )

write.csv(
  df_certain_emotion_candidates,
  here("01Data", "02Derived", "05Manual Inspectation", "Certain_Manual Inspectation.csv"),
  row.names = FALSE
)
```

```{r}

#token level inclusion and exclusion
certain_totals <- amb_utt_certain_flagged %>%
  summarise(
    # token level
    total_certain_tokens = sum(n_certain_tokens, na.rm = TRUE),
    total_certain_emotion_tokens = sum(n_certain_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_certain = sum(n_certain_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_certain= sum(n_certain_emotion > 0, na.rm = TRUE)
  )

certain_totals

#check emotion inclusion examples

set.seed(1)

amb_utt_certain_flagged %>%
  filter(n_certain_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_certain_flagged %>%
  filter(n_certain_tokens > 0, n_certain_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_certain_flagged %>%
  filter(id == 17486037)

```


####3.5 High

```{r}
df_high <- readRDS(here("01Data", "02Derived", "01Dataframe","df_high_utterances.rds"))
tok_high <- readRDS(here("01Data", "02Derived", "02Spacy Parsed","spacy_high_parsed.rds"))
```

Rule: exclude adj+noun

```{r}


# Make sure token_id is numeric
tok_high2 <- tok_high %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id)
  )

high_tok <- tok_high2 %>%
  filter(lemma == "high") %>%
  select(doc_id, token_id, token, lemma, pos, dep_rel)

tok_next <- tok_high2 %>%
  select(doc_id, token_id, next_token = token, next_lemma = lemma, next_pos = pos)

high_tok_with_next <- high_tok %>%
  mutate(next_token_id = token_id + 1L) %>%     # <-- compute next id
  left_join(
    tok_next,
    by = c("doc_id" = "doc_id", "next_token_id" = "token_id")
  ) %>%
  select(-next_token_id)


high_excluded_adj_noun <- high_tok_with_next %>%
  filter(pos == "ADJ", next_pos %in% c("NOUN", "PROPN")) %>%
  distinct(doc_id, token_id)

high_excluded_high_on <- high_tok_with_next %>%
  filter(pos == "ADJ", str_to_lower(next_token) == "on") %>%
  distinct(doc_id, token_id)

high_exclude_start_standalone_docs <- df_high %>%
  filter(
    str_detect(text, "^\\s*high(\\s*[\\.,—\\-]|\\s*$)")
  ) %>%
  distinct(doc_id)
high_excluded_start_standalone <- high_exclude_start_standalone_docs %>%
  left_join(high_tok_with_next %>% distinct(doc_id, token_id), by = "doc_id")

high_excluded_tokens_all <- bind_rows(
  high_excluded_adj_noun,
  high_excluded_high_on,
  high_excluded_start_standalone
  # , high_excluded_high_five
) %>%
  distinct(doc_id, token_id)

# total "high" tokens per utterance
high_all_counts <- high_tok %>%
  count(doc_id, name = "n_high_tokens")

# excluded (non-emotion) "high" tokens per utterance
high_excluded_counts <- high_excluded_tokens_all %>%
  count(doc_id, name = "n_high_nonemotion")

# emotion counts = total - excluded
high_emotion_counts <- high_all_counts %>%
  left_join(high_excluded_counts, by = "doc_id") %>%
  mutate(
    n_high_nonemotion = coalesce(n_high_nonemotion, 0L),
    n_high_emotion = n_high_tokens - n_high_nonemotion
  ) %>%
  select(doc_id, n_high_tokens, n_high_emotion, n_high_nonemotion)

saveRDS(
  high_emotion_counts,
  here("01Data", "02Derived","03Annotation", "high_counts_by_doc.rds")
)
```

Join back to utterance-level dataframe to check

```{r}

high_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","high_counts_by_doc.rds")
)

amb_utt_high_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(high_emotion_counts, by = "doc_id") %>%
  mutate(
    n_high_tokens = coalesce(n_high_tokens, 0L),
    n_high_emotion = coalesce(n_high_emotion, 0L),
    n_high_nonemotion = coalesce(n_high_nonemotion, 0L)
  )

```

```{r}
high_totals <- amb_utt_high_flagged %>%
  summarise(
    # token level
    total_high_tokens = sum(n_high_tokens, na.rm = TRUE),
    total_high_emotion_tokens = sum(n_high_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_high = sum(n_high_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_high = sum(n_high_emotion > 0, na.rm = TRUE)
  )

high_totals


#check emotion inclusion examples

set.seed(1)

amb_utt_high_flagged %>%
  filter(n_high_emotion > 0) %>%
  slice_sample(n = 200) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_high_flagged %>%
  filter(n_high_tokens > 0, n_high_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_high_flagged %>%
  filter(id == 318002)
```

####3.10 Touched

```{r}
df_touched <- readRDS(here("01Data", "02Derived", "01Dataframe","df_touched_utterances.rds"))
tok_touched <- readRDS(here("01Data", "02Derived", "02Spacy Parsed","spacy_touched_parsed.rds"))
```

Rule: exclude adj+noun

```{r}
tok_touched2 <- tok_touched %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id),
    token_lower = str_to_lower(token),
    lemma_lower = str_to_lower(lemma)
  )
touched_tok <- tok_touched2 %>%
  filter(token_lower == "touched") %>%
  select(doc_id, token_id, token, lemma, pos)

# total touched tokens per doc
touched_all_counts <- touched_tok %>%
  count(doc_id, name = "n_touched_tokens")

# excluded = touched used as a VERB (your rule)
touched_excluded_tokens <- touched_tok %>%
  filter(pos == "VERB") %>%
  distinct(doc_id, token_id)

touched_excluded_counts <- touched_excluded_tokens %>%
  count(doc_id, name = "n_touched_nonemotion")

# derive emotion counts
touched_emotion_counts <- touched_all_counts %>%
  left_join(touched_excluded_counts, by = "doc_id") %>%
  mutate(
    n_touched_nonemotion = coalesce(n_touched_nonemotion, 0L),
    n_touched_emotion = n_touched_tokens - n_touched_nonemotion
  ) %>%
  select(doc_id, n_touched_tokens, n_touched_emotion, n_touched_nonemotion)

# save
saveRDS(
  touched_emotion_counts,
  here("01Data", "02Derived","03Annotation", "touched_counts_by_doc.rds")
)

```

Join back to utterance-level dataframe to check

```{r}

touched_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","touched_counts_by_doc.rds")
)

amb_utt_touched_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(touched_emotion_counts, by = "doc_id") %>%
  mutate(
    n_touched_tokens = coalesce(n_touched_tokens, 0L),
    n_touched_emotion = coalesce(n_touched_emotion, 0L),
    n_touched_nonemotion = coalesce(n_touched_nonemotion, 0L)
  )

```

```{r}
touched_totals <- amb_utt_touched_flagged %>%
  summarise(
    # token level
    total_touched_tokens = sum(n_touched_tokens, na.rm = TRUE),
    total_touched_emotion_tokens = sum(n_touched_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_touched = sum(n_touched_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_touched = sum(n_touched_emotion > 0, na.rm = TRUE)
  )

touched_totals


#check emotion inclusion examples

set.seed(1)

amb_utt_touched_flagged %>%
  filter(n_touched_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_touched_flagged %>%
  filter(n_touched_tokens > 0, n_touched_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_touched_flagged %>%
  filter(id == 318002)

touched_tok %>% 
  filter(doc_id == 387228)
```
####3.11 Patient

```{r}
df_patient <- readRDS(here("01Data", "02Derived", "01Dataframe","df_patient_utterances.rds"))
tok_patient <- readRDS(here("01Data", "02Derived", "02Spacy Parsed","spacy_patient_parsed.rds"))
```

Rule: exclude adj+noun

```{r}
tok_patient2 <- tok_patient %>%
  mutate(
    doc_id = as.character(doc_id),
    token_id = as.integer(token_id),
    token_lower = str_to_lower(token),
    lemma_lower = str_to_lower(lemma)
  )
patient_tok <- tok_patient2 %>%
  filter(token_lower == "patient") %>%
  select(doc_id, token_id, token, lemma, pos)

# total patient tokens per doc
patient_all_counts <- patient_tok %>%
  count(doc_id, name = "n_patient_tokens")

# excluded = touched used as a NOUN (your rule)
patient_excluded_tokens <- patient_tok %>%
  filter(pos == "NOUN") %>%
  distinct(doc_id, token_id)

patient_excluded_counts <- patient_excluded_tokens %>%
  count(doc_id, name = "n_patient_nonemotion")

# derive emotion counts
patient_emotion_counts <- patient_all_counts %>%
  left_join(patient_excluded_counts, by = "doc_id") %>%
  mutate(
    n_patient_nonemotion = coalesce(n_patient_nonemotion, 0L),
    n_patient_emotion = n_patient_tokens - n_patient_nonemotion
  ) %>%
  select(doc_id, n_patient_tokens, n_patient_emotion, n_patient_nonemotion)

# save
saveRDS(
patient_emotion_counts,
  here("01Data", "02Derived","03Annotation", "patient_counts_by_doc.rds")
)

```

Join back to utterance-level dataframe to check

```{r}

patient_emotion_counts <- readRDS(
  here("01Data", "02Derived","03Annotation","patient_counts_by_doc.rds")
)

amb_utt_patient_flagged <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(patient_emotion_counts, by = "doc_id") %>%
  mutate(
    n_patient_tokens = coalesce(n_patient_tokens, 0L),
    n_patient_emotion = coalesce(n_patient_emotion, 0L),
    n_patient_nonemotion = coalesce(n_patient_nonemotion, 0L)
  )

```

```{r}
patient_totals <- amb_utt_patient_flagged %>%
  summarise(
    # token level
    total_patient_tokens = sum(n_patient_tokens, na.rm = TRUE),
    total_patient_emotion_tokens = sum(n_patient_emotion, na.rm = TRUE),

    # utterance level
    utterances_with_patient = sum(n_patient_tokens > 0, na.rm = TRUE),
    utterances_with_emotion_patient = sum(n_patient_emotion > 0, na.rm = TRUE)
  )

patient_totals


#check emotion inclusion examples

set.seed(1)

amb_utt_patient_flagged %>%
  filter(n_patient_emotion > 0) %>%
  slice_sample(n = 40) %>%
  select(id, gloss)

#check non-emotions exclusion
amb_utt_patient_flagged %>%
  filter(n_patient_tokens > 0, n_patient_emotion == 0) %>%
 slice_sample(n = 40) %>%
  select(id, gloss)

#sanity check
amb_utt_patient_flagged %>%
  filter(id == 318002)
```

### Step 4:S-BERT Sentence Embedded Tagging

```{r}
library(reticulate)
use_condaenv("spacyr-arm", required = TRUE)
py_install("sentence-transformers", pip = TRUE)
st <- import("sentence_transformers")

amb_utt <- readRDS(here("01Data", "02Derived", "amb_utterances.rds"))

# load the model (this defines `model`)
model <- st$SentenceTransformer("all-MiniLM-L6-v2")
```
#### 4.1 High

prepare High SBERT candidate set
```{r}
high_counts_by_doc <- readRDS(
  here("01Data", "02Derived","03Annotation", "high_counts_by_doc.rds")
)
amb_high_sbert <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(high_counts_by_doc, by = "doc_id") %>%
  filter(!is.na(n_high_emotion) & n_high_emotion > 0) %>% 
  select(
    doc_id,
    text = gloss_lower,
    n_high_emotion,
    n_high_tokens
  )

nrow(amb_high_sbert)                 # how many utterances go to SBERT
summary(amb_high_sbert$n_high_emotion)

# inspect a few
amb_high_sbert %>%
  head(10)
```

```{r}
# --- 0) Data ---
text_col <- "text"
stopifnot(text_col %in% names(amb_high_sbert))

amb_high_sbert2 <- amb_high_sbert %>%
  mutate(.text = str_to_lower(.data[[text_col]])) %>%
  filter(str_detect(.text, "\\bhigh\\b"))

stopifnot(nrow(amb_high_sbert2) > 0)

# --- 1) Sense prototypes (high:excitement or flying high) 

emotion_ref_high <- paste(
  "feeling high",
"excited thrilled euphoric elated",
  "on a high",
  "buzzing energized",
  "intoxicated stoned drunk",
  "in a great mood"
)

other_ref_high <- paste(
  "high up",
  "tall height altitude",
  "far above the ground",
  "flying high",
  "high ceiling",
  "high mountain",
  "high number level value price",
  "high school grade"
)

all_texts <- c(amb_high_sbert2$.text, emotion_ref_high, other_ref_high)

# --- 2) Embeddings (sentence-transformers) ---
# assumes you've already created `model` as you just did:
# model <- st$SentenceTransformer("all-MiniLM-L6-v2")

X <- py_to_r(model$encode(all_texts))
X <- as.matrix(X)
storage.mode(X) <- "double"

stopifnot(nrow(X) == length(all_texts))
stopifnot(ncol(X) > 10)
stopifnot(all(is.finite(X)))

# --- 3) Cosine similarity ---
cos_sim_safe <- function(a, b) {
  denom <- sqrt(sum(a*a)) * sqrt(sum(b*b))
  if (!is.finite(denom) || denom == 0) return(NA_real_)
  sum(a*b) / denom
}

n_u <- nrow(amb_high_sbert2)
i_emotion <- n_u + 1
i_other   <- n_u + 2

sim_emotion <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_emotion, ]), numeric(1))
sim_other   <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_other, ]),   numeric(1))

# --- 4) Results ---
df_high_scored <- amb_high_sbert2 %>%
  mutate(
    sim_to_high_emotion = sim_emotion,
    sim_to_high_other  = sim_other,
   high_sense_margin   = sim_emotion - sim_other,
   high_sense_pred     = ifelse(high_sense_margin > 0, "emotion_like", "other_like")
  )

# Quick summary
mean(df_high_scored$high_sense_pred == "emotion_like", na.rm = TRUE)
summary(df_high_scored$high_sense_margin)

# Inspect extremes
df_high_scored %>% arrange(desc(high_sense_margin)) %>% select(.text, high_sense_margin) %>% head(15) #high closer to emotion
df_high_scored %>% arrange(high_sense_margin) %>% select(.text, high_sense_margin) %>% head(15)

saveRDS(
  df_high_scored,
  here("01Data", "02Derived","04S-BERT Scored", "df_high_scored.rds")
)
```
1. get N-> if > 100, do not manually inspect move on to 2. SBERT -> if mean <0, manually inspect everything >0 -> decide to abandon word or not. 

#### 4.2 Blue
Prepare Blue candidate dataset
```{r}
blue_counts_by_doc <- readRDS(
  here("01Data", "02Derived","03Annotation", "blue_counts_by_doc.rds")
)
amb_blue_sbert <- amb_utt %>%
  mutate(doc_id = as.character(id)) %>%
  left_join(blue_counts_by_doc, by = "doc_id") %>%
  filter(!is.na(n_blue_emotion) & n_blue_emotion > 0) %>% 
  select(
    doc_id,
    text = gloss_lower,
    n_blue_emotion,
    n_blue_tokens
  )

nrow(amb_blue_sbert)                 # how many utterances go to SBERT
summary(amb_blue_sbert$n_blue_emotion)

# inspect a few
amb_blue_sbert %>%
  head(10)
```

```{r}

# --- 0) Data ---
text_col <- "text"
stopifnot(text_col %in% names(amb_blue_sbert))

df_blue2 <- amb_blue_sbert %>%
  mutate(.text = str_to_lower(.data[[text_col]])) %>%
  filter(str_detect(.text, "\\bblue\\b"))

stopifnot(nrow(df_blue2) > 0)

# --- 1) Sense prototypes (blue: sad vs color) ---
emotion_ref_blue <- "feeling blue sad unhappy down depressed lonely upset mood"
color_ref_blue   <- "blue color paint sky shirt eyes crayon visual colour pigment"

all_texts <- c(df_blue2$.text, emotion_ref_blue, color_ref_blue)

# --- 2) Embeddings (sentence-transformers) ---
# assumes you've already created `model` as you just did:
# model <- st$SentenceTransformer("all-MiniLM-L6-v2")

X <- py_to_r(model$encode(all_texts))
X <- as.matrix(X)
storage.mode(X) <- "double"

stopifnot(nrow(X) == length(all_texts))
stopifnot(ncol(X) > 10)
stopifnot(all(is.finite(X)))

# --- 3) Cosine similarity ---
cos_sim_safe <- function(a, b) {
  denom <- sqrt(sum(a*a)) * sqrt(sum(b*b))
  if (!is.finite(denom) || denom == 0) return(NA_real_)
  sum(a*b) / denom
}

n_u <- nrow(df_blue2)
i_emotion <- n_u + 1
i_color   <- n_u + 2

sim_emotion <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_emotion, ]), numeric(1))
sim_color   <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_color, ]),   numeric(1))

# --- 4) Results ---
df_blue_scored <- df_blue2 %>%
  mutate(
    sim_to_blue_emotion = sim_emotion,
    sim_to_blue_color   = sim_color,
    blue_sense_margin   = sim_emotion - sim_color,
    blue_sense_pred     = ifelse(blue_sense_margin > 0, "emotion_like", "color_like")
  )

# Quick summary
mean(df_blue_scored$blue_sense_pred == "emotion_like", na.rm = TRUE)
summary(df_blue_scored$blue_sense_margin)

# Inspect extremes
df_blue_scored %>% arrange(desc(blue_sense_margin)) %>% select(doc_id, .text, blue_sense_margin) %>% head(15) #blue closer to emotion
df_blue_scored %>% arrange(blue_sense_margin) %>% select(.text, blue_sense_margin) %>% head(15)

saveRDS(
  df_blue_scored,
  here("01Data", "02Derived","04S-BERT Scored", "df_blue_scored.rds")
)
```

Maunal Inspectation
```{r}
df_blue_emotion_candidates <- df_blue_scored %>%
  filter(blue_sense_margin > 0) %>% 
 arrange(desc(blue_sense_margin)) %>% 
  select(
    doc_id,
    text = .text,          # or your original text column
    blue_sense_margin
  )

write.csv(
  df_blue_emotion_candidates,
  here("01Data", "02Derived", "05Manual Inspectation", "Blue_Manual Inspectation.csv"),
  row.names = FALSE
)

nrow(df_blue_emotion_candidates)
```


Maunal Inspectation
```{r}
df_blue_emotion_candidates <- df_blue_scored %>%
  filter(blue_sense_margin > 0) %>% 
 arrange(desc(blue_sense_margin)) %>% 
  select(
    doc_id,
    text = .text,          # or your original text column
    blue_sense_margin
  )

write.csv(
  df_blue_emotion_candidates,
  here("01Data", "02Derived", "05Manual Inspectation", "Blue_Manual Inspectation.csv"),
  row.names = FALSE
)

nrow(df_blue_emotion_candidates)
```
#### 4.3 *Lost

```{r}
df_lost <- readRDS(here("01Data", "02Derived", "01Dataframe","df_lost_utterances.rds"))
```

```{r}

# --- 0) Data ---
text_col <- "text"
stopifnot(text_col %in% names(df_lost))

df_lost2 <- df_lost %>%
  mutate(.text = str_to_lower(.data[[text_col]])) %>%
  filter(str_detect(.text, "\\blost\\b"))

stopifnot(nrow(df_lost2) > 0)

# --- 1) Sense prototypes (blue: sad vs color) ---
emotion_ref_lost <- paste(
  "feeling lost",
  "confused unsure uncertain",
  "don't know what to do",
  "worried anxious overwhelmed",
  "helpless",
  "emotionally lost"
)

other_ref_lost <- paste(
  "lost it",
  "lost my toy keys shoes",
  "can't find it",
  "misplaced",
  "lost in the store",
  "not knowing the way",
  "got lost on the road",
  "lost game points"
)

all_texts <- c(df_lost2$.text, emotion_ref_lost, other_ref_lost)

# --- 2) Embeddings (sentence-transformers) ---
# assumes you've already created `model` as you just did:
# model <- st$SentenceTransformer("all-MiniLM-L6-v2")

X <- py_to_r(model$encode(all_texts))
X <- as.matrix(X)
storage.mode(X) <- "double"

stopifnot(nrow(X) == length(all_texts))
stopifnot(ncol(X) > 10)
stopifnot(all(is.finite(X)))

# --- 3) Cosine similarity ---
cos_sim_safe <- function(a, b) {
  denom <- sqrt(sum(a*a)) * sqrt(sum(b*b))
  if (!is.finite(denom) || denom == 0) return(NA_real_)
  sum(a*b) / denom
}

n_u <- nrow(df_lost2)
i_emotion <- n_u + 1
i_other   <- n_u + 2

sim_emotion <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_emotion, ]), numeric(1))
sim_other   <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_other, ]),   numeric(1))

# --- 4) Results ---
df_lost_scored <- df_lost2 %>%
  mutate(
    sim_to_lost_emotion = sim_emotion,
    sim_to_lost_other  = sim_other,
   lost_sense_margin   = sim_emotion - sim_other,
    lost_sense_pred     = ifelse(lost_sense_margin > 0, "emotion_like", "other_like")
  )

# Quick summary
mean(df_lost_scored$lost_sense_pred == "emotion_like", na.rm = TRUE)
summary(df_lost_scored$lost_sense_margin)

# Inspect extremes
df_lost_scored %>% arrange(desc(lost_sense_margin)) %>% select(doc_id, .text, lost_sense_margin) %>% head(15) #lost closer to emotion
df_lost_scored %>% arrange(lost_sense_margin) %>% select(.text, lost_sense_margin) %>% head(15)

saveRDS(
  df_lost_scored,
  here("01Data", "02Derived","04S-BERT Scored", "df_lost_scored.rds")
)
```
manual inspection

```{r}
df_lost_emotion_candidates <- df_lost_scored %>%
  filter(lost_sense_margin > 0) %>% 
 arrange(desc(lost_sense_margin)) %>% 
  select(
    doc_id,
    text = .text,
    lost_sense_margin
  )

write.csv(
  df_lost_emotion_candidates,
  here("01Data", "02Derived", "05Manual Inspectation", "Lost_Manual Inspectation.csv"),
  row.names = FALSE
)

nrow(df_lost_emotion_candidates)
```

#### 3.8 *Quiet

```{r}
df_quiet <- readRDS(here("01Data", "02Derived", "01Dataframe","df_quiet_utterances.rds"))
```

```{r}

# --- 0) Data ---
text_col <- "text"
stopifnot(text_col %in% names(df_quiet))

df_quiet2 <- df_quiet %>%
  mutate(.text = str_to_lower(.data[[text_col]])) %>%
  filter(str_detect(.text, "\\bquiet\\b"))

stopifnot(nrow(df_quiet2) > 0)

# --- 1) Sense prototypes (blue: sad vs color) ---

emotion_ref_quiet <- paste(
  "feeling quiet",
  "withdrawn silent subdued",
  "not talking much",
  "keeping to oneself",
  "low energy",
  "emotionally quiet",
  "not very expressive today"
)

other_ref_quiet <- paste(
  "be quiet",
  "quiet room",
  "quiet voice",
  "no noise",
  "silent environment",
  "quiet place",
  "sound level low"
)

all_texts <- c(df_quiet2$.text, emotion_ref_quiet, other_ref_quiet)

# --- 2) Embeddings (sentence-transformers) ---

X <- py_to_r(model$encode(all_texts))
X <- as.matrix(X)
storage.mode(X) <- "double"

stopifnot(nrow(X) == length(all_texts))
stopifnot(ncol(X) > 10)
stopifnot(all(is.finite(X)))

# --- 3) Cosine similarity ---
cos_sim_safe <- function(a, b) {
  denom <- sqrt(sum(a*a)) * sqrt(sum(b*b))
  if (!is.finite(denom) || denom == 0) return(NA_real_)
  sum(a*b) / denom
}

n_u <- nrow(df_quiet2)
i_emotion <- n_u + 1
i_other   <- n_u + 2

sim_emotion <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_emotion, ]), numeric(1))
sim_other   <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_other, ]),   numeric(1))

# --- 4) Results ---
df_quiet_scored <- df_quiet2 %>%
  mutate(
    sim_to_quiet_emotion = sim_emotion,
    sim_to_quiet_other  = sim_other,
   quiet_sense_margin   = sim_emotion - sim_other,
    quiet_sense_pred     = ifelse(quiet_sense_margin > 0, "emotion_like", "other_like")
  )

# Quick summary
mean(df_quiet_scored$quiet_sense_pred == "emotion_like", na.rm = TRUE)
summary(df_quiet_scored$quiet_sense_margin)

# Inspect extremes
df_quiet_scored %>% arrange(desc(quiet_sense_margin)) %>% select(doc_id, .text, quiet_sense_margin) %>% head(15) #lost closer to emotion
df_quiet_scored %>% arrange(quiet_sense_margin) %>% select(.text, quiet_sense_margin) %>% head(15)

saveRDS(
  df_quiet_scored,
  here("01Data", "02Derived","04S-BERT Scored", "df_quiet_scored.rds")
)
```

```{r}
df_quiet_emotion_candidates <- df_quiet_scored %>%
  filter(quiet_sense_margin > 0) %>% 
 arrange(desc(quiet_sense_margin)) %>% 
  select(
    doc_id,
    text = .text,
    quiet_sense_margin
  )

write.csv(
  df_quiet_emotion_candidates,
  here("01Data", "02Derived", "05Manual Inspectation", "Quiet_Manual Inspectation.csv"),
  row.names = FALSE
)

nrow(df_quiet_emotion_candidates)
```


#### 3.11 *Low

```{r}
df_low <- readRDS(here("01Data", "02Derived", "01Dataframe","df_low_utterances.rds"))
```

```{r}
# --- 0) Data ---
text_col <- "text"
stopifnot(text_col %in% names(df_low))

df_low2 <- df_low %>%
  mutate(.text = str_to_lower(.data[[text_col]])) %>%
  filter(str_detect(.text, "\\blow\\b"))

stopifnot(nrow(df_low2) > 0)

# --- 1) Sense prototypes---

# --- Sense prototypes (low: sadness/energy vs physical level/position) ---

emotion_ref_low <- paste(
  "feeling low",
  "sad unhappy down depressed",
  "low mood",
  "feeling bad",
  "feeling tired drained",
  "low energy"
)

other_ref_low <- paste(
  "low height",
  "near the ground",
  "low table chair",
  "low number level price",
  "low voice sound",
  "low volume",
  "low battery",
  "low score"
)


all_texts <- c(df_low2$.text, emotion_ref_low, other_ref_low)

# --- 2) Embeddings (sentence-transformers) ---

X <- py_to_r(model$encode(all_texts))
X <- as.matrix(X)
storage.mode(X) <- "double"

stopifnot(nrow(X) == length(all_texts))
stopifnot(ncol(X) > 10)
stopifnot(all(is.finite(X)))

# --- 3) Cosine similarity ---
cos_sim_safe <- function(a, b) {
  denom <- sqrt(sum(a*a)) * sqrt(sum(b*b))
  if (!is.finite(denom) || denom == 0) return(NA_real_)
  sum(a*b) / denom
}

n_u <- nrow(df_low2)
i_emotion <- n_u + 1
i_other   <- n_u + 2

sim_emotion <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_emotion, ]), numeric(1))
sim_other   <- vapply(1:n_u, \(i) cos_sim_safe(X[i, ], X[i_other, ]),   numeric(1))

# --- 4) Results ---
df_low_scored <- df_low2 %>%
  mutate(
    sim_to_low_emotion = sim_emotion,
    sim_to_low_other  = sim_other,
   low_sense_margin   = sim_emotion - sim_other,
    low_sense_pred     = ifelse(low_sense_margin > 0, "emotion_like", "other_like")
  )

# Quick summary
mean(df_low_scored$low_sense_pred == "emotion_like", na.rm = TRUE)
summary(df_low_scored$low_sense_margin)

# Inspect extremes
df_low_scored %>% arrange(desc(low_sense_margin)) %>% select(doc_id, .text, low_sense_margin) %>% head(15) #lost closer to emotion
df_low_scored %>% arrange(low_sense_margin) %>% select(.text, low_sense_margin) %>% head(15)

saveRDS(
  df_low_scored,
  here("01Data", "02Derived","04S-BERT Scored", "df_low_scored.rds")
)
```


```{r}
df_low_emotion_candidates <- df_low_scored %>%
  filter(low_sense_margin > 0) %>% 
 arrange(desc(low_sense_margin)) %>% 
  select(
    doc_id,
    text = .text,
    low_sense_margin
  )

write.csv(
  df_low_emotion_candidates,
  here("01Data", "02Derived", "05Manual Inspectation", "Low_Manual Inspectation.csv"),
  row.names = FALSE
)

nrow(df_low_emotion_candidates)
```


### Step 5: Manual Inspectation for the rest of words (counts <100)

```{r}
manual_cols <-c("rotten","alarm","ill","tender","crushed","gloomy","faint","odd")
manual_utt <- utt_partners_emotions %>%
  filter(if_any(all_of(manual_cols), ~ .x > 0)) %>% mutate(
    ambiguous_word = apply(
      select(., all_of(manual_cols)),
      1,
      function(x) {
        paste(manual_cols[x > 0], collapse = "|")
      }
    )
  ) %>% 
  select(ambiguous_word, id, gloss,role_category,gloss_lower)
write.csv(
  manual_utt,
  here("01Data", "02Derived", "05Manual Inspectation", "Manual Inspectation.csv"),
  row.names = FALSE
)
```

All manually checked - to keep 
```{r}
#no emotions: low, high 

blue_keep_ids <- c(
  "17588871","1874322","1874329","2151231","576976","17597709","669881","17665068","577638","726330"
)

lost_keep_ids <- c(
  "1527894","459663","17428246", "2057822","17722692","1259827","1094129","2887378","2818728","2849440","
2007599", "995740","17444311", "17448315","2007599","2584319",
  "17502496","1731805","1091706","1897272","1578728","2007639"
)

certain_keep_ids <- c("371308","545276","1005121","1184301","1333544","1511526","1511761","2842200","2849875","17501783","17662427","7671589","17679994","17686478","17714307","17724249")

alarm_keep_ids <- c("1182355")

crushed_keep_ids <- c("819179","17501726")

faint_keep_ids <- c("452728",
  "554853","643331","666780","698206","946715","946754","1146196","1146203","1954799",
  "1954822","1954834","17271738","17327658"
)

gloomy_keep_ids <- c(
  "1161295","1560492","2242059",
  "17545451","17545507","17545531"
)

ill_keep_ids <- c(
  "673511","749108","1006772","1049666","1071050","1071062",
  "1071078","1071107","1331881","1331893","1331902","1468506","1479532","1488385","1488393","1550136","1951185","2291926","2356751"
)

odd_keep_ids <- c("1266039",
"368924","375262","543852","581460","581510","581543","582494","582616","582957","583214","584559","584734","629928","658670","664948","665055","665068","670460","1438454","1456906","1458011","1477736","1491850","1492446","1504646","1504746","1504828","1507783",
"1507797","1513593","1516144","1520108","1520172","1521941","1526837","1526990","1529600","1534672","1539931","1541834","1545547","1547160","1563228","1563314","1567252","1567909","1567928","1569065","1571522","1897872","2145584","2166605","2357234","2575441","2575454","2575511","2575536","2575565","2575589","2575613","2579033","2602494","2815642","2824508","2826951","2829779","2837529",
"17287014","17438120","17512577","17516705","17516732","17516744","17524422","17524578","17524592","17529103","17529114","17532831","17537369","17537397","17699852","450183","582454","598709","602771","1378593","1545538","1897863","1897891","1921627","2575427","2575524","2575580","2575603","2575683","2575707","17516691","17516671","17529091","17535196","17535208","17535229","17535292","17537384")

rotten_keep_ids <- c("749197","1104561","1476204","1733024","1743442","2042892","2042927","2042961","2052260","2055194","2055386","2055427","2055975","2056072","2056162","2056193","2061308","2065633","2065813","2076216","2076561","2076830","2078829","2079116","2299477","1305291","2055402","2056190")

tender_keep_ids <- c( "1110743","1289804","1339572","1478498","1733271","2134420","2155854","1195584","1339577","2015077","17367220","17382859")
```
###Step 6 Merge back

```{r}

utt_partners_emotions_merged <- utt_partners_emotions %>%
  mutate(
    blue = if_else(id %in% blue_keep_ids, blue, 0L),
    lost = if_else(id %in% lost_keep_ids, lost, 0L),
    certain = if_else(id %in% certain_keep_ids, alarm, 0L),
    alarm = if_else(id %in% alarm_keep_ids, alarm, 0L), 
    crushed = if_else(id %in% crushed_keep_ids, crushed, 0L), 
    faint = if_else(id %in% faint_keep_ids, faint, 0L), 
    gloomy = if_else(id %in% gloomy_keep_ids, gloomy, 0L), 
    ill = if_else(id %in% ill_keep_ids, ill, 0L),
    odd = if_else(id %in% odd_keep_ids, odd, 0L),
    rotten = if_else(id %in% rotten_keep_ids, rotten, 0L),
    tender = if_else(id %in% tender_keep_ids, tender, 0L)
  )

utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
    like_counts_by_doc %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_like_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    like = coalesce(n_like_emotion, 0L)
  ) %>%
  select(-n_like_emotion)

utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
    well_counts_by_doc %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_well_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    well = coalesce(n_well_emotion, 0L)
  ) %>%
  select(-n_well_emotion)

utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
    kind_counts_by_doc %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_kind_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    kind = coalesce(n_kind_emotion, 0L)
  ) %>%
  select(-n_kind_emotion)


utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
    fine_emotion_counts %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_fine_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    fine = coalesce(n_fine_emotion, 0L)
  ) %>%
  select(-n_fine_emotion)


saveRDS(utt_partners_emotions_merged,
        here("01Data","02Derived","utt_partners_emotions_merged.rds"))
```
interim save and clear storage
```{r}
utt_partners_emotions_merged <- readRDS(
  here("01Data","02Derived","utt_partners_emotions_merged.rds")
)

utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
    merry_emotion_counts %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_merry_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    merry = coalesce(n_merry_emotion, 0L)
  ) %>%
  select(-n_merry_emotion)

utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
    moved_emotion_counts %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_moved_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    moved = coalesce(n_moved_emotion, 0L)
  ) %>%
  select(-n_moved_emotion)


utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
   touched_emotion_counts %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_touched_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    touched = coalesce(n_touched_emotion, 0L)
  ) %>%
  select(-n_touched_emotion)

utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(id = as.character(id)) %>%
  left_join(
   patient_emotion_counts %>%
      mutate(doc_id = as.character(doc_id)) %>%
      select(doc_id, n_patient_emotion),
    by = c("id" = "doc_id")
  ) %>%
  mutate(
    patient = coalesce(n_patient_emotion, 0L)
  ) %>%
  select(-n_patient_emotion)

#Exclude high and low
utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(high = 0L) %>% 
  mutate(low = 0L) 
```
#### save output
```{r}
utt_partners_emotions_merged <- utt_partners_emotions_merged %>%
  mutate(
    emotion_count = rowSums(
      pick(all_of(emotion_words)),
      na.rm = TRUE
    )
  )

saveRDS(utt_partners_emotions_merged,
        here("01Data","02Derived","FINAL_utt_RQ1_emotions_merged.rds"))
```

##Sanity Check 
```{r}

utt_partners_emotions_merged %>%
  filter(id == "2822644") %>%
  select(id,gloss, like)
```

Sanity check:

```{r}
id_to_check <- "3778"   # replace with a transcript id

transcript_check <- utt_partners_emotions %>%
  filter(transcript_id == id_to_check) %>%
  select(transcript_id, speaker_code, gloss, emotion_count)

transcript_check <- transcript_check %>%
  filter(emotion_count > 0)

transcript_check
```

Word counts: all partner

```{r}
ut_eng_na_rq1 <- ut_eng_na_rq1 %>%
  mutate(
    word_count = stringr::str_count(gloss, "\\S+")  # counts non-space sequences = words
  )

partner_word_counts <- ut_eng_na_rq1 %>%
  group_by(transcript_id) %>%
  summarise(
    total_partner_words = sum(word_count, na.rm = TRUE),
    total_partner_utterances = n()
  )
```



Exclude: Low
To check: 
Quiet/ Strong - subtle difference 
Pride Rock/ Mock Meek

Questions:

-   would you like one - counted as non-emotion in this version (ADP) as it appeared in a polite preference question

-   how can i be fully confident that I did not exclude other uses of "like" that may broadly refer to emotion?
