---
title: "07RQ1DataAnalysis"
format: html
editor: visual
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set( echo = FALSE, # hide code by default 
                        message = FALSE, # hide messages 
                        warning = FALSE, # hide warnings 
                        results = "hide" # hide printed output from code 
                        )

```

Prep work space and load packages

```{r}
source(here::here("00_setup.R"))
library(lme4)
library(splines)
library(glmmTMB)
```

Load cleaned dataset

```{r}
utt_partners_emotions_merged <- readRDS( here("01Data","02Derived","FINAL_utt_partners_emotions_merged.rds") )

file_lex <- read.csv("~/Desktop/R_WD/ThesisDataAnalysis/1987-Affectivelexicon-foundations_words.csv", stringsAsFactors = FALSE) 
emotion_words <-file_lex$Word
```

## Descriptive Stats Across all transcripts, how many times each words are used + frequency

```{r}

total_word_count <- utt_partners_emotions_merged %>%
  summarise(
    total_words = sum(str_count(gloss_lower, "\\b\\w+\\b"))
  ) %>%
  pull(total_words)

total_emotion_words_merged <- utt_partners_emotions_merged %>%
  summarise(
    across(all_of(emotion_words), ~ sum(.x, na.rm = TRUE)),
   emotion_count = sum(emotion_count, na.rm = TRUE), .groups = "drop"
  )

emotion_counts_long_merged <- total_emotion_words_merged %>%
  pivot_longer(
    cols = everything(),
    names_to = "word",
    values_to = "count"
  ) %>%
  mutate(
    freq_per_1000_words = (count / total_word_count) * 1000
  ) %>%
  arrange(desc(freq_per_1000_words))

emotion_counts_long_merged

write.csv(
  emotion_counts_long_merged,
  "MERGED emotion word count and frequency.csv",
  row.names = FALSE
)
```

By Role
```{r}
total_emotion_words_by_role_merged <- utt_partners_emotions_merged %>%
  group_by(role_category) %>%
  summarise(
    across(all_of(emotion_words), ~ sum(.x, na.rm = TRUE)),
    emotion_count = sum(emotion_count, na.rm = TRUE),
    .groups = "drop"
  )

total_word_count_by_role_merged <- utt_partners_emotions_merged %>%
  mutate(word_count = str_count(gloss_lower, "\\b\\w+\\b")) %>%
  group_by(role_category) %>%
  summarise(
    total_words = sum(word_count, na.rm = TRUE),
    .groups = "drop"
  )

emotion_counts_long_by_role_merged <- total_emotion_words_by_role_merged %>%
  left_join(total_word_count_by_role_merged, by = "role_category") %>%
  pivot_longer(
    cols = -c(role_category, total_words),
    names_to = "word",
    values_to = "count"
  ) %>%
  mutate(
    freq_per_1000_words = (count / total_words) * 1000
  ) %>%
  arrange(role_category, desc(freq_per_1000_words))

total_emotion_counts_long_by_role_merged <- emotion_counts_long_by_role_merged %>%
  filter(
    word == "emotion_count"
  ) %>% 
  select(role_category,count,freq_per_1000_words)
total_emotion_counts_long_by_role_merged

write.csv(
  total_emotion_counts_long_by_role_merged,
  here("03Output", "emotion word count and frequency_by role.csv"),
  row.names = FALSE
)
```
Prep Dataset for RQ1 

```{r}
utt_model_df <- utt_partners_emotions_merged %>%
  select(
    # identifiers
    id,
    target_child_id,
    speaker_id,
    corpus_id,
    transcript_id,
    
    # predictors
    role_category,
    target_child_age,
    target_child_sex,
    
    # outcome / exposure
    emotion_count,
    num_tokens
  )

utt_model_df %>%
count(!is.na(target_child_sex), !is.na(target_child_age))

utt_model_df_rq1 <- utt_model_df %>%
  filter(num_tokens > 0) %>% 
 filter(!is.na(target_child_age), !is.na(target_child_sex))

```
### prep transcript-level dataset
```{r}
transcript_id <- "transcript_id"

df_tr <- utt_model_df_rq1 %>%
  group_by(
    !!sym(transcript_id),
    target_child_id,
    role_category,
    target_child_age,
    target_child_sex
  ) %>%
  summarise(
    emotion_count = sum(emotion_count, na.rm = TRUE),
    num_tokens    = sum(num_tokens, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(num_tokens > 0)

m_tr_nb <- glmmTMB(
  emotion_count ~
    role_category * ns(target_child_age, df = 3) +
    target_child_sex +
    offset(log(num_tokens)) +
    (1 | target_child_id),
  family = nbinom2(),
  data = df_tr
)
m_tr_nb
```

### prediction and plotting
```{r}
# observed age range
age_min_m <- min(df_tr$target_child_age, na.rm = TRUE)
age_max_m <- max(df_tr$target_child_age, na.rm = TRUE)
age_min_y <- age_min_m / 12
age_max_y <- age_max_m / 12

age_years  <- seq(age_min_y, age_max_y, by = 0.1) #sequence of age by 0.1 years 
age_months <- age_years * 12

role_levels <- sort(unique(df_tr$role_category))
sex_levels  <- sort(unique(df_tr$target_child_sex))
ref_sex     <- sex_levels[1]

new_data <- expand.grid( #build grid of all possible combinations 
  target_child_age = age_months,
  role_category = role_levels,
  target_child_sex = ref_sex,
  num_tokens = 1000 #sets offset
)

#convert to factors and match original model
new_data$role_category    <- factor(new_data$role_category, levels = role_levels)
new_data$target_child_sex <- factor(new_data$target_child_sex, levels = sex_levels)
new_data$target_child_age_years <- new_data$target_child_age / 12

pred <- predict(
  m_tr_nb,
  newdata = new_data,
  type = "link",
  se.fit = TRUE,
  re.form = NA
)

#transforming results 
new_data$fit_link <- pred$fit
new_data$se_link  <- pred$se.fit
new_data$fit      <- exp(new_data$fit_link)
new_data$ci_lower <- exp(new_data$fit_link - 1.96 * new_data$se_link)
new_data$ci_upper <- exp(new_data$fit_link + 1.96 * new_data$se_link)

p <- ggplot(new_data,
            aes(x = target_child_age_years, y = fit,
                color = role_category, fill = role_category)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.2, colour = NA) +
  scale_x_continuous(breaks = seq(0, 12, by = 1),
                     limits = c(age_min_y, age_max_y)) +
  coord_cartesian(ylim = c(0, 25)) +
  labs(
    title = "Emotion Words by Partner Role and Child Age",
    x = "Age (years)",
    y = "Expected Count Per 1,000 Words"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5),
    legend.title = element_blank(),
    legend.position = "bottom",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
print(p)

#sanity check

#check for overfitting

# Add the raw rate to your original data frame
df_tr <- df_tr %>%
  mutate(rate_per_1000 = (emotion_count / num_tokens) * 1000)

# Add geom_jitter to your existing plot 'p'
p + geom_jitter(data = df_tr, 
                aes(x = target_child_age/12, y = rate_per_1000, color = role_category),
                alpha = 0.1, size = 0.5)

age_summary <- df_tr %>%
  mutate(age_bin = floor(target_child_age / 12)) %>%
  group_by(age_bin) %>%
  summarise(
    n_transcripts = n(),
    mean_rate_per_1000 = mean(rate_per_1000, na.rm = TRUE),
    median_rate_per_1000 = median(rate_per_1000, na.rm = TRUE),
    max_rate_per_1000 = max(rate_per_1000, na.rm = TRUE)
  )
age_summary
print(p)


summary(utt_model_df_rq1$emotion_count)
quantile(utt_model_df_rq1$emotion_count, c(.9, .95, .99, .999), na.rm = TRUE)
max(utt_model_df_rq1$emotion_count, na.rm = TRUE)
table(utt_model_df_rq1$emotion_count, useNA="no")[1:10]
```





 1000 *
  sum(utt_partners_emotions_merged$emotion_count, na.rm = TRUE) /
  sum(utt_partners_emotions_merged$num_tokens, na.rm = TRUE)
 
 # total tokens that would appear in the DTM
sum(df_child_tr[, word_columns])

# emotion tokens in the same universe
sum(df_child_tr$emotion_words)

 
df_emo_per_tr_all <- utt_child_emotions %>%
  group_by(transcript_id) %>%
  summarise(
    emotion_words_per_transcript = sum(emotion_count, na.rm = TRUE),
    total_tokens = sum(num_tokens, na.rm = TRUE),
    .groups = "drop"
  )

summary(df_emo_per_tr_all$emotion_words_per_transcript)
quantile(df_emo_per_tr_all$emotion_words_per_transcript,
         probs = c(.05, .25, .5, .75, .95))



df_child_rate <- utt_child_emotions %>%
  group_by(transcript_id) %>%
  summarise(
    emo = sum(emotion_count),
    tok = sum(num_tokens),
    rate_per_1000 = 1000 * emo / tok,
    .groups = "drop"
  ) %>%
  filter(tok > 0)

summary(df_child_rate$rate_per_1000)
quantile(df_child_rate$rate_per_1000, c(.25,.5,.75,.9,.95), na.rm=TRUE)
mean(df_child_rate$rate_per_1000, na.rm=TRUE)

```

Test 
Overdispersion was assessed using simulation-based residual diagnostics (DHARMa). Although the dispersion test was statistically significant due to the very large sample size, the dispersion ratio was close to 1 (≈1.04), indicating minimal overdispersion. We therefore report Poisson mixed-effects models with a log offset for utterance length.


```{r}
m_pois <- glmer(
  emotion_count ~
    role_category * target_child_age +
    offset(log(num_tokens)) +
    (1 | target_child_id) +
    (1 | speaker_id),
  family = poisson,
  data = utt_model_df_rq1
)

m_pois

overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp  <- residuals(model, type = "pearson")
  Pearson.chisq <- sum(rp^2)
  ratio <- Pearson.chisq / rdf
  pval  <- pchisq(Pearson.chisq, df = rdf, lower.tail = FALSE)
  c(chisq = Pearson.chisq, ratio = ratio, p = pval)
}

overdisp_fun(m_pois)
```

```{r}
m_null <- glmer(
  emotion_count ~
    offset(log(num_tokens)) +
    (1 | target_child_id) +
    (1 | speaker_id),
  family = poisson,
  data = utt_model_df_rq1
)

m_main_nb <- glmmTMB(
  emotion_count ~
    role_category * ns(target_child_age, df = 4) +
    target_child_sex +
    offset(log(num_tokens)) +
    (1 | target_child_id) +
    (1 | speaker_id),
  family = nbinom2(),
  data = utt_model_df_rq1
)


saveRDS(
  m_main_nb,
  file = "m_main_nb_glmmTMB.rds",
  compress = "xz"
)

m_main_nb <- readRDS("m_main_nb_glmmTMB.rds")

library(ggeffects)
library(ggplot2)


age_grid <- seq(
  from = min(utt_model_df_rq1$target_child_age, na.rm = TRUE),
  to   = max(utt_model_df_rq1$target_child_age, na.rm = TRUE),
  length.out = 100
)
library(emmeans)

emm <- emmeans(
  m_main_nb,
  ~ role_category | target_child_age,
  at = list(
    target_child_age = age_grid,
    num_tokens = 1000   # THIS IS CRITICAL
  ),
  type = "response"
)


emm_df <- as.data.frame(emm)

library(ggplot2)

p <- ggplot(emm_df,
            aes(x = target_child_age, y = response, colour = role_category)) +
  geom_line(size = 1.2) +
  geom_ribbon(aes(ymin = asymp.LCL, ymax = asymp.UCL, fill = role_category),
              alpha = 0.15, colour = NA) +
  coord_cartesian(ylim = c(0, 25)) +
  labs(x="Child age (months)",
       y="Predicted emotion words per 1000 tokens",
       colour="Partner role", fill="Partner role") +
  theme_classic()

p


nrow(utt_model_df_rq1)

sapply(utt_model_df_rq1[c("target_child_id","speaker_id","role_category")], \(x) length(unique(x)))
table(utt_model_df_rq1$emotion_count == 0)
summary(utt_model_df_rq1$num_tokens)

m_main <- glmer(
  emotion_count ~
    role_category * ns(target_child_age, df = 4) +
    target_child_sex +
    offset(log(num_tokens)) +
    (1 | target_child_id) +
    (1 | speaker_id),
  family = poisson,
  data = utt_model_df_rq1
)

library(splines)


m_main
```
```{r}

# ---- 1) Define observed age range (months) and build an age grid in YEARS ----
age_min_m <- min(utt_model_df_rq1$target_child_age, na.rm = TRUE)  # ~ 5.99
age_max_m <- max(utt_model_df_rq1$target_child_age, na.rm = TRUE)  # ~ 144.43

age_min_y <- age_min_m / 12
age_max_y <- age_max_m / 12
age_max_y

age_years  <- seq(age_min_y, age_max_y, by = 0.1)
age_months <- age_years * 12

# ---- 2) Get robust "levels" even if variables are character ----
role_levels <- sort(unique(utt_model_df_rq1$role_category))
sex_levels  <- sort(unique(utt_model_df_rq1$target_child_sex))

# choose a reference sex that exists in your data
# (change this if you want a specific level, e.g., "female")
ref_sex <- sex_levels[1]

# ---- 3) Build prediction grid (set num_tokens=1000 for "per 1,000 words") ----
new_data <- expand.grid(
  target_child_age = age_months,
  role_category = role_levels,
  target_child_sex = ref_sex,
  num_tokens = 1000
)

# Ensure factor coding matches what the model expects
new_data$role_category    <- factor(new_data$role_category, levels = role_levels)
new_data$target_child_sex <- factor(new_data$target_child_sex, levels = sex_levels)

# Add age in years for plotting
new_data <- new_data %>%
  mutate(target_child_age_years = target_child_age / 12)

# ---- 4) Predict population-level (fixed effects only), with SEs ----
pred <- predict(
  m_main_nb,
  newdata = new_data,
  type = "link",     # get link-scale + SE, then transform ourselves
  se.fit = TRUE,
  re.form = NA       # IMPORTANT: exclude random effects for clean trajectories
)

# ---- 5) Build response-scale fit + 95% CI ----
new_data <- new_data %>%
  mutate(
    fit_link = pred$fit,
    se_link  = pred$se.fit,
    ci_lower_link = fit_link - 1.96 * se_link,
    ci_upper_link = fit_link + 1.96 * se_link,
    fit     = exp(fit_link),
    ci_lower = exp(ci_lower_link),
    ci_upper = exp(ci_upper_link)
  )

# Optional: prevent ribbons from extending below 0 (not necessary, but tidy)
new_data <- new_data %>%
  mutate(
    ci_lower = pmax(ci_lower, 0)
  )

# ---- 6) Plot (match your supervisor-style: line + ribbon, x in years, y 0–25) ----
p_roles <- ggplot(
  new_data,
  aes(x = target_child_age_years, y = fit, color = role_category, fill = role_category)
) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = ci_lower, ymax = ci_upper), alpha = 0.2, colour = NA) +
  scale_x_continuous(
    breaks = seq(floor(age_min_y), ceiling(age_max_y), by = 1),
    limits = c(age_min_y, age_max_y)
  ) +
  coord_cartesian(ylim = c(0, 25)) +
  labs(
    title = "Emotion Words by Partner Role and Child Age",
    x = "Age (years)",
    y = "Expected Count Per 1,000 Words"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
    axis.title.x = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    axis.ticks = element_line(color = "black"),
    legend.title = element_blank(),
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(1, 1, 1, 1, "cm")
  )

print(p_roles)
```



```{r}
audit_my <- ut_eng_na_rq1 %>%
  mutate(gloss_lower = tolower(gloss)) %>%
  filter(transcript_id %in% tr_ids) %>%
  summarise(
    emo_tokens = sum(gloss_lower %in% emotion_words, na.rm = TRUE),
    total_tokens = n(),
    .by = transcript_id
  )
audit_my
audit_wide <- utt_partners_emotions %>%   # the wide table you bind_cols'd onto
  filter(transcript_id %in% tr_ids) %>%
  summarise(
    emo_tokens_wide = rowSums(across(all_of(emotion_words)), na.rm = TRUE),
    .by = transcript_id
  )
audit_wide
audit_wide
  audit_my %>%
  inner_join(audit_wide, by = "transcript_id") %>%
  mutate(diff = emo_tokens - emo_tokens_wide)

check <- utt_partners_emotions_merged %>% 
filter(transcript_id == "3676")
check
```
  