---
title: "10 – Additional Analysis: Partner Emotion Word Use as Predictor of Child Emotion Word Use"
format: html
editor: visual
---

# Overview

**Research question** — If social partners talk more about emotion, do children talk more about emotion too? And does this association change across development?

This analysis operates at the **transcript level** (one observation per transcript), complementing the turn-by-turn CRQA approach in script 09.

Two sections are reported:

**Part 1 — Full sample**: All transcripts retained. All non-child speakers collapsed into one "partner" category. Tests the overall partner → child association and its developmental moderation and mediation.

**Part 2 — Partner-type comparison**: Restricted to transcripts where the child speaks to exactly one partner type. Tests whether the partner → child association differs by who the child is talking to (Mother, Father, Sibling, etc.).

------------------------------------------------------------------------

## 0. Setup

```{r setup, message=FALSE, warning=FALSE}
source(here::here("00_setup.R"))
library(lme4)
library(lmerTest)
library(splines)
library(emmeans)
library(MuMIn)        # partial R² via nested model comparison
library(broom.mixed)
library(ggeffects)
```

------------------------------------------------------------------------

## 1. Load data

```{r load}
utt <- readRDS(here("01Data", "02Derived", "FINAL_utt_RQ1_emotions_merged.rds"))

cat("Rows:        ", nrow(utt), "\n")
cat("Transcripts: ", n_distinct(utt$transcript_id), "\n")
cat("Children:    ", n_distinct(utt$target_child_id), "\n")
```

------------------------------------------------------------------------

## 2. Build transcript-level age lookup

`target_child_age` is only populated on child-speaker rows in CHILDES. Build a reliable transcript-level lookup from child utterances only.

```{r age-lookup}
age_lookup <- utt %>%
  filter(role_category == "Target Child", !is.na(target_child_age)) %>%
  group_by(transcript_id) %>%
  summarise(
    target_child_age = first(target_child_age),
    target_child_sex = first(target_child_sex),
    target_child_id  = first(target_child_id),
    corpus_name      = first(corpus_name),
    .groups = "drop"
  )
```

------------------------------------------------------------------------

# PART 1 — Full Sample (all partner types collapsed)

------------------------------------------------------------------------

## 3. Aggregate all transcripts — partner as single category

All non-child utterances are summed into one "partner" series per transcript, regardless of who is speaking.

```{r part1-aggregate}
speaker_summary_all <- utt %>%
  mutate(speaker_category = if_else(role_category == "Target Child",
                                    "child", "partner")) %>%
  group_by(transcript_id, speaker_category) %>%
  summarise(
    total_tokens  = sum(num_tokens,    na.rm = TRUE),
    total_emotion = sum(emotion_count, na.rm = TRUE),
    n_utterances  = n(),
    .groups = "drop"
  ) %>%
  mutate(emo_rate = if_else(total_tokens > 0,
                             total_emotion / total_tokens, NA_real_))

transcript_all <- speaker_summary_all %>%
  pivot_wider(
    id_cols     = transcript_id,
    names_from  = speaker_category,
    values_from = c(emo_rate, total_tokens, total_emotion, n_utterances)
  ) %>%
  rename(
    child_emo_rate   = emo_rate_child,
    partner_emo_rate = emo_rate_partner,
    child_tokens     = total_tokens_child,
    partner_tokens   = total_tokens_partner,
    child_emotion    = total_emotion_child,
    partner_emotion  = total_emotion_partner,
    child_utts       = n_utterances_child,
    partner_utts     = n_utterances_partner
  ) %>%
  left_join(age_lookup, by = "transcript_id") %>%
  mutate(target_child_age_yrs = target_child_age / 12)

# --- Stepwise attrition: shows which filter removes the most children ---
# (transcript_all_pre holds all columns before any filter)
transcript_all_pre <- transcript_all  # at this point filters not yet applied

attrition_table <- tibble(
  Step = c(
    "0. Pre-filter (all transcripts in utt)",
    "1. + child & partner both have tokens > 0",
    "2. + age data present",
    "3. + child >= 10 tokens",
    "4. + partner >= 10 tokens  [final]"
  ),
  n_transcripts = c(
    nrow(transcript_all_pre),
    nrow(transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0)),
    nrow(transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0, !is.na(target_child_age_yrs))),
    nrow(transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0, !is.na(target_child_age_yrs), child_tokens >= 10)),
    nrow(transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0, !is.na(target_child_age_yrs), child_tokens >= 10, partner_tokens >= 10))
  ),
  n_children = c(
    n_distinct(transcript_all_pre$target_child_id),
    n_distinct((transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0))$target_child_id),
    n_distinct((transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0, !is.na(target_child_age_yrs)))$target_child_id),
    n_distinct((transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0, !is.na(target_child_age_yrs), child_tokens >= 10))$target_child_id),
    n_distinct((transcript_all_pre %>% filter(!is.na(child_tokens), child_tokens > 0, !is.na(partner_tokens), partner_tokens > 0, !is.na(target_child_age_yrs), child_tokens >= 10, partner_tokens >= 10))$target_child_id)
  )
) %>%
  mutate(transcripts_lost = lag(n_transcripts, default = first(n_transcripts)) - n_transcripts,
         children_lost    = lag(n_children,    default = first(n_children))    - n_children)

print(attrition_table)

# Apply all filters
transcript_all <- transcript_all %>%
  filter(
    !is.na(child_emo_rate),
    !is.na(partner_emo_rate),
    !is.na(target_child_age_yrs),
    child_tokens   >= 10,
    partner_tokens >= 10
  )

cat("\nFinal analytic sample:\n")
cat("Transcripts:", nrow(transcript_all), "\n")
cat("Children:   ", n_distinct(transcript_all$target_child_id), "\n")
cat("Age range:  ", round(min(transcript_all$target_child_age_yrs), 2),
    "–", round(max(transcript_all$target_child_age_yrs), 2), "years\n")
```

------------------------------------------------------------------------

## 4. Descriptives — full sample

```{r part1-desc}
transcript_all %>%
  summarise(
    n_transcripts    = n(),
    n_children       = n_distinct(target_child_id),
    age_min          = round(min(target_child_age_yrs),  2),
    age_max          = round(max(target_child_age_yrs),  2),
    child_emo_mean   = round(mean(child_emo_rate),       4),
    child_emo_sd     = round(sd(child_emo_rate),         4),
    partner_emo_mean = round(mean(partner_emo_rate),     4),
    partner_emo_sd   = round(sd(partner_emo_rate),       4)
  ) %>% print()

cor.test(transcript_all$child_emo_rate, transcript_all$partner_emo_rate)
```

------------------------------------------------------------------------

## 5. Within-child / between-child decomposition

```{r part1-decompose}
transcript_all <- transcript_all %>%
  group_by(target_child_id) %>%
  mutate(
    partner_emo_between = mean(partner_emo_rate, na.rm = TRUE),
    partner_emo_within  = partner_emo_rate - partner_emo_between
  ) %>%
  ungroup()
```

------------------------------------------------------------------------

## 6. Moderation — does age moderate the partner → child association?

### 6a. Fit model

```{r part1-mod-model}
m_mod <- lmer(
  child_emo_rate ~
    partner_emo_within  * ns(target_child_age_yrs, df = 3) +
    partner_emo_between * ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_all,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_mod)
```

### 6b. Test within × age interaction

```{r part1-mod-anova}
m_mod_noint <- lmer(
  child_emo_rate ~
    partner_emo_within +
    partner_emo_between * ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_all,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

anova(m_mod_noint, update(m_mod, REML = FALSE))
```

### 6c. Simple slopes — partner effect at each age

```{r part1-simple-slopes}
emtrends(m_mod,
         var   = "partner_emo_within",
         specs = ~ target_child_age_yrs,
         at    = list(target_child_age_yrs = c(2, 4, 6, 8, 10)))
```

### 6d. Moderation plot

```{r part1-mod-plot}
pred_mod <- ggpredict(
  m_mod,
  terms = c("partner_emo_within [all]", "target_child_age_yrs [2, 4, 6, 8, 10]")
)

ggplot(pred_mod, aes(x = x, y = predicted,
                     colour = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  labs(
    title    = "Partner → child emotion rate by child age (full sample)",
    subtitle = "Within-child effect  |  marginalised over sex and between-child rate",
    x        = "Partner emotion rate (within-child deviation)",
    y        = "Predicted child emotion rate",
    colour   = "Child age (years)", fill = "Child age (years)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

> **Note (Section 6d):** The moderation plot above is generated directly from `m_mod`, which uses `ns(target_child_age_yrs, df = 3)` — a natural spline — not binned or categorical age. Predictions are drawn from the fitted spline model via `ggpredict()`.

### 6e. Partial R² — unique variance explained by key predictors

Semi-partial R² for each term is estimated by nested model comparison: marginal R² of the full model minus marginal R² of a model with that term removed (Nakagawa & Schielzeth, 2013; computed via `MuMIn::r.squaredGLMM()`). Models are refit with ML (REML = FALSE) for comparability.

```{r part1-partial-r2}
# Refit models with ML for nested comparison
m_mod_ml <- update(m_mod, REML = FALSE)

# Without partner_emo_within (main effect + interaction removed; random slope simplified)
m_no_within_ml <- lmer(
  child_emo_rate ~
    ns(target_child_age_yrs, df = 3) +
    partner_emo_between * ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 | target_child_id),
  data    = transcript_all,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

# Without partner_emo_within × age interaction (main effect retained) — m_mod_noint exists
m_no_int_ml <- update(m_mod_noint, REML = FALSE)

# Without partner_emo_between (main effect + interaction removed)
m_no_between_ml <- lmer(
  child_emo_rate ~
    partner_emo_within * ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_all,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

r2_full       <- r.squaredGLMM(m_mod_ml)[1,       "R2m"]
r2_no_within  <- r.squaredGLMM(m_no_within_ml)[1,  "R2m"]
r2_no_int     <- r.squaredGLMM(m_no_int_ml)[1,     "R2m"]
r2_no_between <- r.squaredGLMM(m_no_between_ml)[1, "R2m"]

partial_r2_table <- tibble(
  Term = c(
    "partner_emo_within (main effect + age interaction)",
    "partner_emo_within × age interaction only",
    "partner_emo_between (main effect + age interaction)"
  ),
  Full_R2m    = round(r2_full, 4),
  Reduced_R2m = round(c(r2_no_within, r2_no_int, r2_no_between), 4),
  Partial_R2  = round(c(
    r2_full - r2_no_within,
    r2_full - r2_no_int,
    r2_full - r2_no_between
  ), 4)
)

print(partial_r2_table)
```

### 6f. Corpus random effect — is accounting for CHILDES corpus necessary?

CHILDES transcripts come from different corpora (labs, studies, demographic groups) that may systematically differ in emotion word rates. We assess this in two steps: (1) compute the corpus-level intraclass correlation (ICC) from an unconditional model, and (2) attempt to add corpus as a random effect to the main model. Non-convergence when adding `(1 | corpus_name)` typically signals near-zero corpus variance — which itself justifies omitting it.

```{r part1-corpus-re}
# Step 1: corpus-level ICC from unconditional means model
m_icc_corpus <- lmer(
  child_emo_rate ~ 1 + (1 | corpus_name) + (1 | target_child_id),
  data = transcript_all, REML = TRUE
)
vc_icc <- as.data.frame(VarCorr(m_icc_corpus))
corpus_var <- vc_icc$vcov[vc_icc$grp == "corpus_name"]
child_var  <- vc_icc$vcov[vc_icc$grp == "target_child_id"]
resid_var  <- vc_icc$vcov[vc_icc$grp == "Residual"]
icc_corpus <- corpus_var / (corpus_var + child_var + resid_var)
cat(sprintf("Corpus-level ICC = %.4f  (%.1f%% of total variance)\n",
            icc_corpus, icc_corpus * 100))
cat("Rule of thumb: ICC < .05 = corpus clustering negligible.\n\n")

# Step 2: descriptive means per corpus
corpus_desc <- transcript_all %>%
  group_by(corpus_name) %>%
  summarise(n = n(), n_children = n_distinct(target_child_id),
            mean_child_emo = round(mean(child_emo_rate, na.rm = TRUE), 5),
            .groups = "drop") %>%
  arrange(desc(n))
print(corpus_desc)

# Step 3: add corpus RE to main model — non-convergence confirms near-zero variance
m_mod_corpus <- suppressWarnings(
  lmer(
    child_emo_rate ~
      partner_emo_within  * ns(target_child_age_yrs, df = 3) +
      partner_emo_between * ns(target_child_age_yrs, df = 3) +
      target_child_sex +
      (1 + partner_emo_within | target_child_id) +
      (1 | corpus_name),
    data    = transcript_all,
    REML    = TRUE,
    control = lmerControl(optimizer = "nloptwrap",
                          optCtrl   = list(algorithm = "NLOPT_LN_BOBYQA",
                                           maxeval   = 2e5))
  )
)

vc_corpus <- as.data.frame(VarCorr(m_mod_corpus))
corpus_pct <- vc_corpus$vcov[vc_corpus$grp == "corpus_name"] /
              sum(vc_corpus$vcov) * 100
cat(sprintf("\nCorpus variance in full model = %.1f%% of total\n", corpus_pct))
cat("Variance components:\n")
print(vc_corpus)
```

**Decision:** If corpus ICC < .05 and corpus variance < 5% of total, the base model `m_mod` is well-justified. Non-convergence is consistent with near-zero corpus variance and further supports this conclusion. Report in the methods as: *"Corpus-level clustering was examined via ICC and a random-intercept model; corpus variance accounted for < X% of total variance (ICC = .XX), so corpus was not included as a random effect."*

------------------------------------------------------------------------

## 8. Visualise child and partner rates across age — full sample

```{r part1-age-plot}
rates_by_age <- transcript_all %>%
  mutate(age_block = floor(target_child_age_yrs)) %>%
  filter(age_block <= 11) %>%
  group_by(age_block) %>%
  summarise(
    n            = n(),
    child_mean   = mean(child_emo_rate,   na.rm = TRUE),
    child_se     = sd(child_emo_rate,     na.rm = TRUE) / sqrt(n()),
    partner_mean = mean(partner_emo_rate, na.rm = TRUE),
    partner_se   = sd(partner_emo_rate,   na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_longer(c(child_mean, partner_mean),
               names_to = "speaker", values_to = "emo_rate") %>%
  mutate(
    se      = if_else(speaker == "child_mean", child_se, partner_se),
    speaker = if_else(speaker == "child_mean", "Child", "Partner")
  )

ggplot(rates_by_age, aes(x = age_block, y = emo_rate,
                          colour = speaker, fill = speaker)) +
  geom_ribbon(aes(ymin = emo_rate - se, ymax = emo_rate + se),
              alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 0:11) +
  labs(
    title    = "Child and partner emotion word rate across development (full sample)",
    subtitle = "Mean ± SE per 1-year age block",
    x = "Child age (years)", y = "Emotion word rate (count / tokens)",
    colour = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

------------------------------------------------------------------------

# PART 2 — Partner-type comparison (dyadic transcripts only)

Only transcripts where the target child speaks to exactly **one** partner type are retained so that each transcript has an unambiguous partner category label.

------------------------------------------------------------------------

## 9. Filter to dyadic transcripts

```{r part2-dyadic}
dyadic_info <- utt %>%
  filter(role_category != "Target Child") %>%
  group_by(transcript_id) %>%
  summarise(
    n_partner_types = n_distinct(role_category),
    partner_type    = if_else(n_distinct(role_category) == 1,
                              first(role_category), NA_character_),
    .groups = "drop"
  )

dyadic_ids <- dyadic_info %>% filter(n_partner_types == 1)

cat("Total transcripts:            ", n_distinct(utt$transcript_id), "\n")
cat("Dyadic (1 partner type):      ", nrow(dyadic_ids), "\n")
cat("Excluded (multiple partners): ",
    n_distinct(utt$transcript_id) - nrow(dyadic_ids), "\n\n")

cat("Transcripts per partner category:\n")
dyadic_ids %>% count(partner_type, name = "n") %>% arrange(desc(n)) %>% print()
```

------------------------------------------------------------------------

## 10. Aggregate dyadic transcripts

```{r part2-aggregate}
utt_dyadic <- utt %>%
  filter(transcript_id %in% dyadic_ids$transcript_id) %>%
  left_join(dyadic_ids %>% dplyr::select(transcript_id, partner_type),
            by = "transcript_id")

speaker_summary_dyadic <- utt_dyadic %>%
  mutate(speaker_category = if_else(role_category == "Target Child",
                                    "child", "partner")) %>%
  group_by(transcript_id, speaker_category) %>%
  summarise(
    partner_type  = first(partner_type),
    total_tokens  = sum(num_tokens,    na.rm = TRUE),
    total_emotion = sum(emotion_count, na.rm = TRUE),
    n_utterances  = n(),
    .groups = "drop"
  ) %>%
  mutate(emo_rate = if_else(total_tokens > 0,
                             total_emotion / total_tokens, NA_real_))

transcript_dyadic <- speaker_summary_dyadic %>%
  pivot_wider(
    id_cols     = c(transcript_id, partner_type),
    names_from  = speaker_category,
    values_from = c(emo_rate, total_tokens, total_emotion, n_utterances)
  ) %>%
  rename(
    child_emo_rate   = emo_rate_child,
    partner_emo_rate = emo_rate_partner,
    child_tokens     = total_tokens_child,
    partner_tokens   = total_tokens_partner,
    child_emotion    = total_emotion_child,
    partner_emotion  = total_emotion_partner,
    child_utts       = n_utterances_child,
    partner_utts     = n_utterances_partner
  ) %>%
  left_join(age_lookup, by = "transcript_id") %>%
  mutate(target_child_age_yrs = target_child_age / 12) %>%
  filter(
    !is.na(child_emo_rate),
    !is.na(partner_emo_rate),
    !is.na(target_child_age_yrs),
    child_tokens   >= 10,
    partner_tokens >= 10
  ) %>%
  mutate(
    partner_type = factor(partner_type),
    partner_type = relevel(partner_type, ref = "Mother")
  )

# Drop partner types with fewer than 20 transcripts
keep_types <- transcript_dyadic %>%
  count(partner_type) %>%
  filter(n >= 20) %>%
  pull(partner_type)

transcript_dyadic <- transcript_dyadic %>% filter(partner_type %in% keep_types)

# Within / between decomposition
transcript_dyadic <- transcript_dyadic %>%
  group_by(target_child_id) %>%
  mutate(
    partner_emo_between = mean(partner_emo_rate, na.rm = TRUE),
    partner_emo_within  = partner_emo_rate - partner_emo_between
  ) %>%
  ungroup()

cat("Dyadic transcripts retained:", nrow(transcript_dyadic), "\n")
cat("Children:                   ", n_distinct(transcript_dyadic$target_child_id), "\n")
cat("\nFinal n per partner type:\n")
print(count(transcript_dyadic, partner_type))
```

------------------------------------------------------------------------

## 11. Descriptives by partner type

```{r part2-desc}
transcript_dyadic %>%
  group_by(partner_type) %>%
  summarise(
    n                = n(),
    age_mean         = round(mean(target_child_age_yrs), 2),
    child_emo_mean   = round(mean(child_emo_rate,   na.rm = TRUE), 4),
    child_emo_sd     = round(sd(child_emo_rate,     na.rm = TRUE), 4),
    partner_emo_mean = round(mean(partner_emo_rate, na.rm = TRUE), 4),
    partner_emo_sd   = round(sd(partner_emo_rate,   na.rm = TRUE), 4),
    .groups = "drop"
  ) %>%
  arrange(desc(n)) %>%
  print()
```

------------------------------------------------------------------------

## 12. Does child emotion rate differ by partner type?

```{r part2-level-model}
m_partner_level <- lmer(
  child_emo_rate ~
    partner_type +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 | target_child_id),
  data    = transcript_dyadic,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_partner_level)

# Estimated marginal means
emm_partner <- emmeans(m_partner_level, ~ partner_type) %>% as_tibble()

ggplot(emm_partner,
       aes(x = reorder(partner_type, emmean), y = emmean,
           colour = partner_type)) +
  geom_hline(yintercept = mean(emm_partner$emmean),
             linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL),
                  linewidth = 0.8, fatten = 3) +
  coord_flip() +
  labs(
    title    = "Child emotion word rate by partner type",
    subtitle = "Model-estimated marginal means  |  adjusted for age and sex  |  95% CI\nDashed = grand mean",
    x = NULL, y = "Predicted child emotion rate"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

## 13. Does the partner → child responsiveness differ by partner type?

```{r part2-slope-model}
m_partner_slope <- lmer(
  child_emo_rate ~
    partner_emo_within * partner_type +
    partner_emo_between +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_dyadic,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_partner_slope)

# Model comparison — does the interaction improve fit?
m_partner_slope_noint <- lmer(
  child_emo_rate ~
    partner_emo_within + partner_type +
    partner_emo_between +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_dyadic,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

anova(m_partner_slope_noint, update(m_partner_slope, REML = FALSE))
```

------------------------------------------------------------------------

## 14. Partner → child slopes per partner type

```{r part2-slopes}
slope_est <- emtrends(m_partner_slope,
                      var   = "partner_emo_within",
                      specs = ~ partner_type) %>%
  as_tibble() %>%
  arrange(desc(partner_emo_within.trend)) %>%
  mutate(sig = if_else(lower.CL > 0 | upper.CL < 0, "sig", "n.s."))

print(slope_est)

ggplot(slope_est,
       aes(x = reorder(partner_type, partner_emo_within.trend),
           y = partner_emo_within.trend, colour = sig)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL),
                  linewidth = 0.8, fatten = 3) +
  scale_colour_manual(values = c("sig" = "#c0392b", "n.s." = "grey60")) +
  coord_flip() +
  labs(
    title    = "Partner → child emotion word slope by partner type",
    subtitle = "Within-child effect  |  95% CI  |  red = CI excludes zero",
    x = NULL, y = "Slope (partner_emo_within → child_emo_rate)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

## 14b. Moderation by partner type — does age moderate the partner → child slope separately for each partner category?

First, check how many transcripts and children are available per partner type. Partner types with fewer transcripts will produce less reliable age-moderation estimates.

```{r part2-mod-feasibility}
n_by_type <- transcript_dyadic %>%
  group_by(partner_type) %>%
  summarise(
    n_transcripts = n(),
    n_children    = n_distinct(target_child_id),
    age_min       = round(min(target_child_age_yrs),  1),
    age_max       = round(max(target_child_age_yrs),  1),
    .groups = "drop"
  ) %>%
  arrange(desc(n_transcripts))

print(n_by_type)

# Eligible: enough transcripts for a spline moderation + random effects
# Conservative threshold: >= 100 transcripts AND >= 20 children
eligible_mod_types <- n_by_type %>%
  filter(n_transcripts >= 100, n_children >= 20) %>%
  pull(partner_type)

cat("\nEligible for per-type moderation model:", paste(eligible_mod_types, collapse = ", "), "\n")
```

For each eligible partner type, fit the same spline moderation model as Section 6 and collect predictions. Partner types below the threshold are excluded from this plot.

```{r part2-mod-by-type}
pred_list <- map(eligible_mod_types, function(pt) {
  d_pt <- transcript_dyadic %>%
    filter(partner_type == pt) %>%
    group_by(target_child_id) %>%
    mutate(
      partner_emo_between = mean(partner_emo_rate, na.rm = TRUE),
      partner_emo_within  = partner_emo_rate - partner_emo_between
    ) %>%
    ungroup()

  m_pt <- tryCatch(
    lmer(
      child_emo_rate ~
        partner_emo_within * ns(target_child_age_yrs, df = 3) +
        partner_emo_between +
        target_child_sex +
        (1 + partner_emo_within | target_child_id),
      data    = d_pt,
      REML    = TRUE,
      control = lmerControl(optimizer = "bobyqa")
    ),
    error = function(e) {
      # Fallback: random intercept only if random slope fails to converge
      lmer(
        child_emo_rate ~
          partner_emo_within * ns(target_child_age_yrs, df = 3) +
          partner_emo_between +
          target_child_sex +
          (1 | target_child_id),
        data    = d_pt,
        REML    = TRUE,
        control = lmerControl(optimizer = "bobyqa")
      )
    }
  )

  ggpredict(
    m_pt,
    terms = c("partner_emo_within [all]",
              "target_child_age_yrs [2, 4, 6, 8, 10]")
  ) %>%
    as_tibble() %>%
    mutate(partner_type = pt)
})

pred_by_type <- bind_rows(pred_list)

ggplot(pred_by_type, aes(x = x, y = predicted,
                          colour = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.8) +
  facet_wrap(~ partner_type) +
  labs(
    title    = "Partner → child emotion rate by age, per partner type",
    subtitle = "Within-child effect  |  per-type spline moderation models  |  marginalised over sex and between-child rate",
    x        = "Partner emotion rate (within-child deviation)",
    y        = "Predicted child emotion rate",
    colour   = "Child age (years)", fill = "Child age (years)"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top",
        strip.text = element_text(face = "bold"))
```

------------------------------------------------------------------------

## 14c. Three-way interaction — does the age moderation of the partner → child slope differ by partner type? (optional)

Tests whether the developmental trajectory of the within-child partner effect varies across partner categories. This extends `m_partner_slope` (which estimates a single age-by-partner_type moderation) to a full three-way interaction.

```{r part2-three-way}
m_three_way <- lmer(
  child_emo_rate ~
    partner_emo_within * ns(target_child_age_yrs, df = 3) * partner_type +
    partner_emo_between +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_dyadic,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_three_way)

# LRT: does the three-way interaction improve fit over the two-way partner × age model?
m_base_ml <- lmer(
  child_emo_rate ~
    partner_emo_within * partner_type +
    ns(target_child_age_yrs, df = 3) +
    partner_emo_between +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_dyadic,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

m_three_way_ml <- update(m_three_way, REML = FALSE)

anova(m_base_ml, m_three_way_ml)
```

------------------------------------------------------------------------

## 15b. Subsample representativeness check

Compares the analytic sample (`transcript_all`, after applying token filters) against the excluded transcripts (in `utt` but not retained) on child age distribution, sex ratio, and emotion word rates.

```{r subsample-comparison}
# Build full-census transcript-level frame from utt (before token filters)
transcript_census <- utt %>%
  mutate(speaker_category = if_else(role_category == "Target Child", "child", "partner")) %>%
  group_by(transcript_id, speaker_category) %>%
  summarise(
    total_tokens  = sum(num_tokens,    na.rm = TRUE),
    total_emotion = sum(emotion_count, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(emo_rate = if_else(total_tokens > 0, total_emotion / total_tokens, NA_real_)) %>%
  pivot_wider(
    id_cols     = transcript_id,
    names_from  = speaker_category,
    values_from = c(emo_rate, total_tokens)
  ) %>%
  rename(
    child_emo_rate   = emo_rate_child,
    partner_emo_rate = emo_rate_partner,
    child_tokens     = total_tokens_child,
    partner_tokens   = total_tokens_partner
  ) %>%
  left_join(age_lookup, by = "transcript_id") %>%
  mutate(target_child_age_yrs = target_child_age / 12) %>%
  filter(!is.na(target_child_age_yrs), !is.na(child_emo_rate), !is.na(partner_emo_rate)) %>%
  mutate(in_sample = transcript_id %in% transcript_all$transcript_id)

cat("Full census transcripts: ", nrow(transcript_census), "\n")
cat("Analytic sample:         ", sum(transcript_census$in_sample), "\n")
cat("Excluded:                ", sum(!transcript_census$in_sample), "\n\n")

# 1. Age distribution
age_in  <- transcript_census %>% filter( in_sample) %>% pull(target_child_age_yrs)
age_out <- transcript_census %>% filter(!in_sample) %>% pull(target_child_age_yrs)
age_t   <- t.test(age_in, age_out)
age_ks  <- ks.test(age_in, age_out)
cat(sprintf("Age (years) — In: M = %.2f (SD = %.2f) | Out: M = %.2f (SD = %.2f)\n",
            mean(age_in), sd(age_in), mean(age_out), sd(age_out)))
cat(sprintf("  t-test: t = %.3f, df = %.1f, p = %.4f\n",
            age_t$statistic, age_t$parameter, age_t$p.value))
cat(sprintf("  KS test: D = %.3f, p = %.4f\n\n", age_ks$statistic, age_ks$p.value))

# 2. Sex ratio
sex_contingency <- table(
  in_sample = transcript_census$in_sample,
  sex       = transcript_census$target_child_sex
)
print(sex_contingency)
sex_chi <- chisq.test(sex_contingency)
cat(sprintf("\nSex χ²(%d) = %.3f, p = %.4f\n\n",
            sex_chi$parameter, sex_chi$statistic, sex_chi$p.value))

# 3. Emotion word rates
emo_child_in  <- transcript_census %>% filter( in_sample) %>% pull(child_emo_rate)
emo_child_out <- transcript_census %>% filter(!in_sample) %>% pull(child_emo_rate)
emo_part_in   <- transcript_census %>% filter( in_sample) %>% pull(partner_emo_rate)
emo_part_out  <- transcript_census %>% filter(!in_sample) %>% pull(partner_emo_rate)

ct_child <- t.test(emo_child_in,  emo_child_out)
ct_part  <- t.test(emo_part_in,   emo_part_out)

cat(sprintf("Child emo rate — In: %.5f | Out: %.5f | t = %.3f, p = %.4f\n",
            mean(emo_child_in, na.rm=TRUE), mean(emo_child_out, na.rm=TRUE),
            ct_child$statistic, ct_child$p.value))
cat(sprintf("Partner emo rate — In: %.5f | Out: %.5f | t = %.3f, p = %.4f\n",
            mean(emo_part_in, na.rm=TRUE), mean(emo_part_out, na.rm=TRUE),
            ct_part$statistic, ct_part$p.value))
```

------------------------------------------------------------------------

## 15. Save

```{r save}
saveRDS(transcript_all,     here("01Data", "02Derived", "additional_full_sample.rds"))
saveRDS(transcript_dyadic,  here("01Data", "02Derived", "additional_dyadic.rds"))
saveRDS(m_mod,              here("01Data", "02Derived", "additional_mod_model.rds"))
saveRDS(m_partner_level,    here("01Data", "02Derived", "additional_partner_level_model.rds"))
saveRDS(m_partner_slope,    here("01Data", "02Derived", "additional_partner_slope_model.rds"))
saveRDS(m_three_way,        here("01Data", "02Derived", "additional_three_way_model.rds"))
```
