---
title: "10 – Additional Analysis: Partner Emotion Word Use as Predictor of Child Emotion Word Use"
format: html
editor: visual
---

# Overview

**Research question** — If social partners talk more about emotion, do children talk more about emotion too? And does this association change across development?

This analysis operates at the **transcript level** (one observation per transcript), complementing the turn-by-turn CRQA approach in script 09.

Two sections are reported:

**Part 1 — Full sample**: All transcripts retained. All non-child speakers collapsed into one "partner" category. Tests the overall partner → child association and its developmental moderation and mediation.

**Part 2 — Partner-type comparison**: Restricted to transcripts where the child speaks to exactly one partner type. Tests whether the partner → child association differs by who the child is talking to (Mother, Father, Sibling, etc.).

------------------------------------------------------------------------

## 0. Setup

```{r setup, message=FALSE, warning=FALSE}
source(here::here("00_setup.R"))
library(lme4)
library(lmerTest)
library(splines)
library(emmeans)
library(mediation)
library(broom.mixed)
library(ggeffects)
```

------------------------------------------------------------------------

## 1. Load data

```{r load}
utt <- readRDS(here("01Data", "02Derived", "FINAL_utt_RQ1_emotions_merged.rds"))

cat("Rows:        ", nrow(utt), "\n")
cat("Transcripts: ", n_distinct(utt$transcript_id), "\n")
cat("Children:    ", n_distinct(utt$target_child_id), "\n")
```

------------------------------------------------------------------------

## 2. Build transcript-level age lookup

`target_child_age` is only populated on child-speaker rows in CHILDES. Build a reliable transcript-level lookup from child utterances only.

```{r age-lookup}
age_lookup <- utt %>%
  filter(role_category == "Target Child", !is.na(target_child_age)) %>%
  group_by(transcript_id) %>%
  summarise(
    target_child_age = first(target_child_age),
    target_child_sex = first(target_child_sex),
    target_child_id  = first(target_child_id),
    .groups = "drop"
  )
```

------------------------------------------------------------------------

# PART 1 — Full Sample (all partner types collapsed)

------------------------------------------------------------------------

## 3. Aggregate all transcripts — partner as single category

All non-child utterances are summed into one "partner" series per transcript, regardless of who is speaking.

```{r part1-aggregate}
speaker_summary_all <- utt %>%
  mutate(speaker_category = if_else(role_category == "Target Child",
                                    "child", "partner")) %>%
  group_by(transcript_id, speaker_category) %>%
  summarise(
    total_tokens  = sum(num_tokens,    na.rm = TRUE),
    total_emotion = sum(emotion_count, na.rm = TRUE),
    n_utterances  = n(),
    .groups = "drop"
  ) %>%
  mutate(emo_rate = if_else(total_tokens > 0,
                             total_emotion / total_tokens, NA_real_))

transcript_all <- speaker_summary_all %>%
  pivot_wider(
    id_cols     = transcript_id,
    names_from  = speaker_category,
    values_from = c(emo_rate, total_tokens, total_emotion, n_utterances)
  ) %>%
  rename(
    child_emo_rate   = emo_rate_child,
    partner_emo_rate = emo_rate_partner,
    child_tokens     = total_tokens_child,
    partner_tokens   = total_tokens_partner,
    child_emotion    = total_emotion_child,
    partner_emotion  = total_emotion_partner,
    child_utts       = n_utterances_child,
    partner_utts     = n_utterances_partner
  ) %>%
  left_join(age_lookup, by = "transcript_id") %>%
  mutate(target_child_age_yrs = target_child_age / 12) %>%
  filter(
    !is.na(child_emo_rate),
    !is.na(partner_emo_rate),
    !is.na(target_child_age_yrs),
    child_tokens   >= 10,
    partner_tokens >= 10
  )

cat("Transcripts (full sample):", nrow(transcript_all), "\n")
cat("Children:                 ", n_distinct(transcript_all$target_child_id), "\n")
cat("Age range:                ", round(min(transcript_all$target_child_age_yrs), 2),
    "–", round(max(transcript_all$target_child_age_yrs), 2), "years\n")
```

------------------------------------------------------------------------

## 4. Descriptives — full sample

```{r part1-desc}
transcript_all %>%
  summarise(
    n_transcripts    = n(),
    n_children       = n_distinct(target_child_id),
    age_min          = round(min(target_child_age_yrs),  2),
    age_max          = round(max(target_child_age_yrs),  2),
    child_emo_mean   = round(mean(child_emo_rate),       4),
    child_emo_sd     = round(sd(child_emo_rate),         4),
    partner_emo_mean = round(mean(partner_emo_rate),     4),
    partner_emo_sd   = round(sd(partner_emo_rate),       4)
  ) %>% print()

cor.test(transcript_all$child_emo_rate, transcript_all$partner_emo_rate)
```

------------------------------------------------------------------------

## 5. Within-child / between-child decomposition

```{r part1-decompose}
transcript_all <- transcript_all %>%
  group_by(target_child_id) %>%
  mutate(
    partner_emo_between = mean(partner_emo_rate, na.rm = TRUE),
    partner_emo_within  = partner_emo_rate - partner_emo_between
  ) %>%
  ungroup()
```

------------------------------------------------------------------------

## 6. Moderation — does age moderate the partner → child association?

### 6a. Fit model

```{r part1-mod-model}
m_mod <- lmer(
  child_emo_rate ~
    partner_emo_within  * ns(target_child_age_yrs, df = 3) +
    partner_emo_between * ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_all,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_mod)
```

### 6b. Test within × age interaction

```{r part1-mod-anova}
m_mod_noint <- lmer(
  child_emo_rate ~
    partner_emo_within +
    partner_emo_between * ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_all,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

anova(m_mod_noint, update(m_mod, REML = FALSE))
```

### 6c. Simple slopes — partner effect at each age

```{r part1-simple-slopes}
emtrends(m_mod,
         var   = "partner_emo_within",
         specs = ~ target_child_age_yrs,
         at    = list(target_child_age_yrs = c(2, 4, 6, 8, 10)))
```

### 6d. Moderation plot

```{r part1-mod-plot}
pred_mod <- ggpredict(
  m_mod,
  terms = c("partner_emo_within [all]", "target_child_age_yrs [2, 4, 6, 8, 10]")
)

ggplot(pred_mod, aes(x = x, y = predicted,
                     colour = group, fill = group)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  labs(
    title    = "Partner → child emotion rate by child age (full sample)",
    subtitle = "Within-child effect  |  marginalised over sex and between-child rate",
    x        = "Partner emotion rate (within-child deviation)",
    y        = "Predicted child emotion rate",
    colour   = "Child age (years)", fill = "Child age (years)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

------------------------------------------------------------------------

## 7. Mediation — does partner rate explain the age effect on child rate?

**Causal chain**: Age → partner_emo_rate → child_emo_rate

### 7a. Path a: age → partner rate

```{r part1-path-a}
m_med_a <- lmer(
  partner_emo_rate ~
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 | target_child_id),
  data = transcript_all,
  REML = TRUE
)
summary(m_med_a)
```

### 7b. Path b: partner rate → child rate (controlling age)

```{r part1-path-b}
m_med_b <- lmer(
  child_emo_rate ~
    partner_emo_rate +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 | target_child_id),
  data = transcript_all,
  REML = TRUE
)
summary(m_med_b)
```

### 7c. Bootstrap indirect effect

```{r part1-mediation}
med_data <- transcript_all %>%
  filter(!is.na(target_child_age_yrs), !is.na(partner_emo_rate),
         !is.na(child_emo_rate), !is.na(target_child_sex))

m_med_a_lm <- lm(partner_emo_rate ~ target_child_age_yrs + target_child_sex,
                  data = med_data)

m_med_b_lm <- lm(child_emo_rate ~ partner_emo_rate + target_child_age_yrs +
                    target_child_sex, data = med_data)

set.seed(42)
med_out <- mediate(
  model.m  = m_med_a_lm,
  model.y  = m_med_b_lm,
  treat    = "target_child_age_yrs",
  mediator = "partner_emo_rate",
  boot     = TRUE,
  sims     = 1000
)

summary(med_out)
plot(med_out)
```

------------------------------------------------------------------------

## 8. Visualise child and partner rates across age — full sample

```{r part1-age-plot}
rates_by_age <- transcript_all %>%
  mutate(age_block = floor(target_child_age_yrs)) %>%
  filter(age_block <= 11) %>%
  group_by(age_block) %>%
  summarise(
    n            = n(),
    child_mean   = mean(child_emo_rate,   na.rm = TRUE),
    child_se     = sd(child_emo_rate,     na.rm = TRUE) / sqrt(n()),
    partner_mean = mean(partner_emo_rate, na.rm = TRUE),
    partner_se   = sd(partner_emo_rate,   na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  ) %>%
  pivot_longer(c(child_mean, partner_mean),
               names_to = "speaker", values_to = "emo_rate") %>%
  mutate(
    se      = if_else(speaker == "child_mean", child_se, partner_se),
    speaker = if_else(speaker == "child_mean", "Child", "Partner")
  )

ggplot(rates_by_age, aes(x = age_block, y = emo_rate,
                          colour = speaker, fill = speaker)) +
  geom_ribbon(aes(ymin = emo_rate - se, ymax = emo_rate + se),
              alpha = 0.2, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 0:11) +
  labs(
    title    = "Child and partner emotion word rate across development (full sample)",
    subtitle = "Mean ± SE per 1-year age block",
    x = "Child age (years)", y = "Emotion word rate (count / tokens)",
    colour = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```

------------------------------------------------------------------------

# PART 2 — Partner-type comparison (dyadic transcripts only)

Only transcripts where the target child speaks to exactly **one** partner type are retained so that each transcript has an unambiguous partner category label.

------------------------------------------------------------------------

## 9. Filter to dyadic transcripts

```{r part2-dyadic}
dyadic_info <- utt %>%
  filter(role_category != "Target Child") %>%
  group_by(transcript_id) %>%
  summarise(
    n_partner_types = n_distinct(role_category),
    partner_type    = if_else(n_distinct(role_category) == 1,
                              first(role_category), NA_character_),
    .groups = "drop"
  )

dyadic_ids <- dyadic_info %>% filter(n_partner_types == 1)

cat("Total transcripts:            ", n_distinct(utt$transcript_id), "\n")
cat("Dyadic (1 partner type):      ", nrow(dyadic_ids), "\n")
cat("Excluded (multiple partners): ",
    n_distinct(utt$transcript_id) - nrow(dyadic_ids), "\n\n")

cat("Transcripts per partner category:\n")
dyadic_ids %>% count(partner_type, name = "n") %>% arrange(desc(n)) %>% print()
```

------------------------------------------------------------------------

## 10. Aggregate dyadic transcripts

```{r part2-aggregate}
utt_dyadic <- utt %>%
  filter(transcript_id %in% dyadic_ids$transcript_id) %>%
  left_join(dyadic_ids %>% dplyr::select(transcript_id, partner_type),
            by = "transcript_id")

speaker_summary_dyadic <- utt_dyadic %>%
  mutate(speaker_category = if_else(role_category == "Target Child",
                                    "child", "partner")) %>%
  group_by(transcript_id, speaker_category) %>%
  summarise(
    partner_type  = first(partner_type),
    total_tokens  = sum(num_tokens,    na.rm = TRUE),
    total_emotion = sum(emotion_count, na.rm = TRUE),
    n_utterances  = n(),
    .groups = "drop"
  ) %>%
  mutate(emo_rate = if_else(total_tokens > 0,
                             total_emotion / total_tokens, NA_real_))

transcript_dyadic <- speaker_summary_dyadic %>%
  pivot_wider(
    id_cols     = c(transcript_id, partner_type),
    names_from  = speaker_category,
    values_from = c(emo_rate, total_tokens, total_emotion, n_utterances)
  ) %>%
  rename(
    child_emo_rate   = emo_rate_child,
    partner_emo_rate = emo_rate_partner,
    child_tokens     = total_tokens_child,
    partner_tokens   = total_tokens_partner,
    child_emotion    = total_emotion_child,
    partner_emotion  = total_emotion_partner,
    child_utts       = n_utterances_child,
    partner_utts     = n_utterances_partner
  ) %>%
  left_join(age_lookup, by = "transcript_id") %>%
  mutate(target_child_age_yrs = target_child_age / 12) %>%
  filter(
    !is.na(child_emo_rate),
    !is.na(partner_emo_rate),
    !is.na(target_child_age_yrs),
    child_tokens   >= 10,
    partner_tokens >= 10
  ) %>%
  mutate(
    partner_type = factor(partner_type),
    partner_type = relevel(partner_type, ref = "Mother")
  )

# Drop partner types with fewer than 20 transcripts
keep_types <- transcript_dyadic %>%
  count(partner_type) %>%
  filter(n >= 20) %>%
  pull(partner_type)

transcript_dyadic <- transcript_dyadic %>% filter(partner_type %in% keep_types)

# Within / between decomposition
transcript_dyadic <- transcript_dyadic %>%
  group_by(target_child_id) %>%
  mutate(
    partner_emo_between = mean(partner_emo_rate, na.rm = TRUE),
    partner_emo_within  = partner_emo_rate - partner_emo_between
  ) %>%
  ungroup()

cat("Dyadic transcripts retained:", nrow(transcript_dyadic), "\n")
cat("Children:                   ", n_distinct(transcript_dyadic$target_child_id), "\n")
cat("\nFinal n per partner type:\n")
print(count(transcript_dyadic, partner_type))
```

------------------------------------------------------------------------

## 11. Descriptives by partner type

```{r part2-desc}
transcript_dyadic %>%
  group_by(partner_type) %>%
  summarise(
    n                = n(),
    age_mean         = round(mean(target_child_age_yrs), 2),
    child_emo_mean   = round(mean(child_emo_rate,   na.rm = TRUE), 4),
    child_emo_sd     = round(sd(child_emo_rate,     na.rm = TRUE), 4),
    partner_emo_mean = round(mean(partner_emo_rate, na.rm = TRUE), 4),
    partner_emo_sd   = round(sd(partner_emo_rate,   na.rm = TRUE), 4),
    .groups = "drop"
  ) %>%
  arrange(desc(n)) %>%
  print()
```

------------------------------------------------------------------------

## 12. Does child emotion rate differ by partner type?

```{r part2-level-model}
m_partner_level <- lmer(
  child_emo_rate ~
    partner_type +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 | target_child_id),
  data    = transcript_dyadic,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_partner_level)

# Estimated marginal means
emm_partner <- emmeans(m_partner_level, ~ partner_type) %>% as_tibble()

ggplot(emm_partner,
       aes(x = reorder(partner_type, emmean), y = emmean,
           colour = partner_type)) +
  geom_hline(yintercept = mean(emm_partner$emmean),
             linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL),
                  linewidth = 0.8, fatten = 3) +
  coord_flip() +
  labs(
    title    = "Child emotion word rate by partner type",
    subtitle = "Model-estimated marginal means  |  adjusted for age and sex  |  95% CI\nDashed = grand mean",
    x = NULL, y = "Predicted child emotion rate"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

## 13. Does the partner → child responsiveness differ by partner type?

```{r part2-slope-model}
m_partner_slope <- lmer(
  child_emo_rate ~
    partner_emo_within * partner_type +
    partner_emo_between +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_dyadic,
  REML    = TRUE,
  control = lmerControl(optimizer = "bobyqa")
)

summary(m_partner_slope)

# Model comparison — does the interaction improve fit?
m_partner_slope_noint <- lmer(
  child_emo_rate ~
    partner_emo_within + partner_type +
    partner_emo_between +
    ns(target_child_age_yrs, df = 3) +
    target_child_sex +
    (1 + partner_emo_within | target_child_id),
  data    = transcript_dyadic,
  REML    = FALSE,
  control = lmerControl(optimizer = "bobyqa")
)

anova(m_partner_slope_noint, update(m_partner_slope, REML = FALSE))
```

------------------------------------------------------------------------

## 14. Partner → child slopes per partner type

```{r part2-slopes}
slope_est <- emtrends(m_partner_slope,
                      var   = "partner_emo_within",
                      specs = ~ partner_type) %>%
  as_tibble() %>%
  arrange(desc(partner_emo_within.trend)) %>%
  mutate(sig = if_else(lower.CL > 0 | upper.CL < 0, "sig", "n.s."))

print(slope_est)

ggplot(slope_est,
       aes(x = reorder(partner_type, partner_emo_within.trend),
           y = partner_emo_within.trend, colour = sig)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_pointrange(aes(ymin = lower.CL, ymax = upper.CL),
                  linewidth = 0.8, fatten = 3) +
  scale_colour_manual(values = c("sig" = "#c0392b", "n.s." = "grey60")) +
  coord_flip() +
  labs(
    title    = "Partner → child emotion word slope by partner type",
    subtitle = "Within-child effect  |  95% CI  |  red = CI excludes zero",
    x = NULL, y = "Slope (partner_emo_within → child_emo_rate)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

------------------------------------------------------------------------

## 15. Save

```{r save}
saveRDS(transcript_all,     here("01Data", "02Derived", "additional_full_sample.rds"))
saveRDS(transcript_dyadic,  here("01Data", "02Derived", "additional_dyadic.rds"))
saveRDS(m_mod,              here("01Data", "02Derived", "additional_mod_model.rds"))
saveRDS(m_partner_level,    here("01Data", "02Derived", "additional_partner_level_model.rds"))
saveRDS(m_partner_slope,    here("01Data", "02Derived", "additional_partner_slope_model.rds"))
```
